---
title: "Children Explanation Seeking Analysis Rmd"
output:
  pdf_document: default
html_document: default
date: "2023-06-13"
---
  ## Libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lme4)
library(tidyverse)
library(summarytools)
library(glmmTMB)

source("./functions/diagnostic_fcns.r")
source("./functions/glmm_stability.r")
source("./functions/boot_glmm.r")
source("./functions/boot_glmm2.r")
source("./functions/glmmTMB_stability.r")
source("./functions/drop1_para_glmmtmb.r")
source("./functions/extract_ranef_gmmTMB.r")

```
#Original data
## Data preparation
Read in the individual data files
```{r}
#Extract file names
list_csv_files <- list.files(path = "./data/children_individual_data_files/", 
               pattern = "*.xlsx", 
               full.names = T)
library("readxl")
test<-read_excel("./data/children_individual_data_files/ES_Child_101.xlsx")
orig.data <- NULL
#Read in all csv files and add the file name as variable
for(f in list_csv_files){  
  ReadInFile <- read_excel(path=f)%>%rename("file_name"=1)
  ReadInFile$child_file <- f
  orig.data <- rbind(orig.data, ReadInFile)
}

orig.data<-orig.data %>%
  rename(subject = Subject...7,
         condition = Subject...11,
         trial_order = `Trial Order`,
         behavioral_category = `Behavioral category`,
         start = `Start (s)`,
         stop= `Stop (s)`,
         duration= `Duration (s)`) %>%
  mutate(start = as.integer(start))%>%
  arrange(subject, start) %>%
  mutate(behavioral_category = fct_recode(as.factor(behavioral_category), "Time spent" = "Time Spent", "Researcher status" = "Researcher Status"))%>%
  mutate(Behavior = fct_recode(as.factor(Behavior), "Researcher absent" = "Researcher Absent", "Researcher present" = "Researcher Present"))%>%
  filter(!(stop=="235.969"& Behavior=="Right box (side)"& subject=="ES101"), !(duration=="4.749"& Behavior=="Right box (side)"& subject=="ES102"))

levels(orig.data$behavioral_category)
levels(as.factor(orig.data$Behavior))
```
data checks
```{r}

n.subj<-length(levels(as.factor(orig.data$subject)))
levels(as.factor(orig.data$child_file))

#add trial number
trial_data <- orig.data %>%
  filter(behavioral_category == "Trial type")

table(trial_data$subject, trial_data$FPS)

trial_data$Trial = rep(c(1:4), n.subj*2)
trial_data <- trial_data %>% select(subject, condition, behavioral_category, start, Trial)

table(trial_data$subject, trial_data$Trial)
levels(as.factor(orig.data$behavioral_category))

#add researcher status
researcher_status_data <- orig.data %>%
  filter(behavioral_category == "Researcher status")%>%
  mutate(researcher_status = Behavior) %>%
  select(subject, condition, researcher_status, behavioral_category, start)%>%
  droplevels()
table(researcher_status_data$subject, researcher_status_data$researcher_status)

#add peeking interest period 
peeking_IP_data <- orig.data %>%
  filter(Behavior == "Toy to end of trial" | Behavior == "Functional" | Behavior == "Non-functional")%>%
  mutate(IP = Behavior) %>%
  select(subject, condition, IP,start)%>%
  droplevels()

table(peeking_IP_data$subject, peeking_IP_data$IP)


orig.data2<-orig.data%>%
  full_join(trial_data)%>%
  fill(Trial) %>%
  full_join(researcher_status_data)%>%
  fill(researcher_status) %>%
  full_join(peeking_IP_data)%>%
  fill(IP)%>%
  mutate(Block = as.numeric(ifelse(
    trial_order == 1 & condition == "Functional",
    1,
    ifelse(
      trial_order == 2 & condition == "Functional",
      2,
      ifelse(
        trial_order == 1 & condition == "Non-functional",
        2,
        ifelse(trial_order == 2 &
                 condition == "Non-functional", 1, "")
      )
    )
  )))
levels(as.factor(orig.data2$Behavior))
orig.data2 %>% filter(Behavior == "Asking why") 
#happened only three times in two subjects

table(orig.data2$behavioral_category, orig.data2$Behavior)
table(orig.data2$researcher_status, orig.data2$Behavior)
table(orig.data2$subject, orig.data2$Trial)
```

### Calculate duration of each trial
```{r}
trialdurationdata <- orig.data2 %>%
  filter(Behavior == "Toy to end of trial") %>% #time chunk in which peeking should happen # old "Researcher absent"
  rename(Trial_duration = duration) %>%
  select(subject, condition, Block, Trial,  trial_order, Trial_duration)%>%
  mutate(Trial_duration = as.numeric(Trial_duration))#%>%
  # filter(!(subject=="ES117" & condition == "Non-functional" & Trial == 3))%>%#filter out two trials because the experimenter absent duration is shorter (<15 sec) or longer than expected (>40 sec)
  # filter(!(subject=="ES135" & condition == "Non-functional" & Trial == 4))

hist(trialdurationdata$Trial_duration)
mean(trialdurationdata$Trial_duration)
min(trialdurationdata$Trial_duration)

table(trialdurationdata$subject,trialdurationdata$condition)
trialdurationdata %>% group_by(condition)%>% summarise(mean_trial_duration = mean(Trial_duration), sd_trial_duration = sd(Trial_duration), se_trial_duration = sd(Trial_duration)/sqrt(length(Trial_duration)))
```

### peeking data
create dataframe just with the peeking categories and add the zeros and the trial duration

```{r}
all_combinations <-
  expand.grid(
    subject = levels(as.factor(orig.data2$subject)),
    condition = levels(as.factor(orig.data2$condition)),
    Trial = as.numeric(levels(as.factor(orig.data2$Trial))),
    Behavior = c("Peek on top", "Peephole peek", "Right side peek", "Right box (side)")
  )

table(orig.data2$IP, orig.data2$subject)
```

# Interest period: toy to end of trial
```{r}
peeking_data <- orig.data2 %>%
  #filter(researcher_status == "Researcher absent") %>%
  filter(IP == "Toy to end of trial") %>%
  filter(Behavior %in% c("Peek on top", "Peephole peek", "Right side peek", "Right box (side)")) %>%
  mutate(duration = as.numeric(duration))%>%
  group_by(subject, condition, trial_order, Block, Trial, Behavior) %>%
  summarise(sum_duration = sum(duration), count_peeking = sum(duration>0))%>%
  ungroup()%>%
  select(-trial_order, -Block) %>%
  full_join((all_combinations)) %>%
  inner_join(trialdurationdata, by = c("subject", "condition", "Trial")) %>%
  complete(fill=list(sum_duration=0, count_peeking=0)) %>%
  mutate(trial_overall = as.numeric(ifelse(Block == 1, Trial,
                                           ifelse(Block == 2, Trial+4, ""))),
         proportion_peeking = sum_duration / Trial_duration)




table(peeking_data$subject, peeking_data$Trial)

write.csv(peeking_data, file = "data/child_explanation_seeking_peeking_data.csv")



```


### Plot of mean peeking
```{r}
peeking_data_agg<-peeking_data %>%
  group_by(subject, condition, Behavior) %>%
  summarise(mean.peeking = mean(sum_duration), count = sum(count_peeking))

peeking_data_agg_trial<-peeking_data %>%
  group_by(subject, condition, Trial, Behavior) %>%
  summarise(mean.peeking = mean(sum_duration), count = sum(count_peeking), proportion_peeking = sum(proportion_peeking))

peeking_data_agg_trial2<-peeking_data_agg_trial %>%
  group_by(subject, condition, Trial) %>%
  summarise(peeking_count_response = sum(count), peeking_response_binary  = sum(peeking_count_response > 0))

peeking_data_proportion<-peeking_data %>%
  group_by(subject, condition, Trial) %>%
  summarise(prop_peeking = sum(proportion_peeking), count = sum(sum_duration > 0))
```

# Reli data
## Data preparation
Read in the individual data files
```{r}
#Extract file names
list_csv_files <- list.files(path = "./data/children_individual_data_files/reli_data", 
               pattern = "*.xlsx", 
               full.names = T)
library("readxl")

reli.raw.data <- NULL
#Read in all csv files and add the file name as variable
for(f in list_csv_files){  
  ReadInFile <- read_excel(path=f)%>%rename("file_name"=1)
  ReadInFile$child_file <- f
  reli.raw.data <- rbind(reli.raw.data, ReadInFile)
}

reli.raw.data<-reli.raw.data %>%
  mutate(file_name = str_remove(file_name, "_Child_")) %>%
  mutate(file_name = str_remove(file_name, "_RC"))%>%
  rename(subject = file_name,
         condition = Subject...11,
         trial_order = `Trial Order`,
         behavioral_category = `Behavioral category`,
         start = `Start (s)`,
         stop= `Stop (s)`,
         duration= `Duration (s)`) %>%
  mutate(start = as.integer(start))%>%
  arrange(subject, start) %>%
  mutate(behavioral_category = fct_recode(as.factor(behavioral_category), "Time spent" = "Time Spent", "Researcher status" = "Researcher Status"))%>%
  mutate(Behavior = fct_recode(as.factor(Behavior), "Researcher absent" = "Researcher Absent", "Researcher present" = "Researcher Present"))


levels(reli.raw.data$behavioral_category)
levels(as.factor(reli.raw.data$Behavior))
```
data checks
```{r}

n.subj<-length(levels(as.factor(reli.raw.data$subject)))
levels(as.factor(reli.raw.data$child_file))

#add trial number
trial_data_reli <- reli.raw.data %>%
  filter(behavioral_category == "Trial type")%>%
    arrange(subject, start) 

table(trial_data_reli$subject, trial_data_reli$FPS)

trial_data_reli$Trial = rep(c(1:4), n.subj*2)
trial_data_reli <- trial_data_reli %>% select(subject, condition, behavioral_category, start, Trial)

table(trial_data_reli$subject, trial_data_reli$Trial)
levels(as.factor(reli.raw.data$behavioral_category))

#add researcher status
researcher_status_data_reli <- reli.raw.data %>%
  filter(behavioral_category == "Researcher status")%>%
  mutate(researcher_status = Behavior) %>%
  select(subject, condition, researcher_status, behavioral_category, start)%>%
  droplevels()
table(researcher_status_data_reli$subject, researcher_status_data_reli$researcher_status)

#add peeking interest period 
peeking_IP_data_reli <- reli.raw.data %>%
  filter(Behavior == "Toy to end of trial" | Behavior == "Functional" | Behavior == "Non-functional")%>%
  mutate(IP = Behavior) %>%
  select(subject, condition, IP,start)%>%
  droplevels()

table(peeking_IP_data_reli$subject, peeking_IP_data_reli$IP)


reli.raw.data2<-reli.raw.data%>%
  full_join(trial_data_reli)%>%
  fill(Trial) %>%
  full_join(researcher_status_data_reli)%>%
  fill(researcher_status) %>%
  full_join(peeking_IP_data_reli)%>%
  fill(IP)%>%
  mutate(Block = as.numeric(ifelse(
    trial_order == 1 & condition == "Functional",
    1,
    ifelse(
      trial_order == 2 & condition == "Functional",
      2,
      ifelse(
        trial_order == 1 & condition == "Non-functional",
        2,
        ifelse(trial_order == 2 &
                 condition == "Non-functional", 1, "")
      )
    )
  )))
levels(as.factor(reli.raw.data2$Behavior))
reli.raw.data2 %>% filter(Behavior == "Asking why") 
#happened only three times in two subjects

table(reli.raw.data2$behavioral_category, reli.raw.data2$Behavior)
table(reli.raw.data2$researcher_status, reli.raw.data2$Behavior)
table(reli.raw.data2$subject, reli.raw.data2$Trial)

table(trial_data_reli$subject, trial_data_reli$condition, trial_data_reli$Trial)
```

### Calculate duration of each trial
```{r}
trialdurationdata_reli <- reli.raw.data2 %>%
  filter(Behavior == "Toy to end of trial") %>% #time chunk in which peeking should happen # old "Researcher absent"
  rename(Trial_duration = duration) %>%
  select(subject, condition, Block, Trial,  trial_order, Trial_duration)%>%
  mutate(Trial_duration = as.numeric(Trial_duration))#%>%
  # filter(!(subject=="ES117" & condition == "Non-functional" & Trial == 3))%>%#filter out two trials because the experimenter absent duration is shorter (<15 sec) or longer than expected (>40 sec)
  # filter(!(subject=="ES135" & condition == "Non-functional" & Trial == 4))

hist(trialdurationdata_reli$Trial_duration)
mean(trialdurationdata_reli$Trial_duration)
min(trialdurationdata_reli$Trial_duration)

table(trialdurationdata_reli$subject,trialdurationdata_reli$condition)
trialdurationdata_reli %>% group_by(condition)%>% summarise(mean_trial_duration = mean(Trial_duration), sd_trial_duration = sd(Trial_duration), se_trial_duration = sd(Trial_duration)/sqrt(length(Trial_duration)))
```

### peeking data
create dataframe just with the peeking categories and add the zeros and the trial duration

```{r}
all_combinations_reli <-
  expand.grid(
    subject = levels(as.factor(reli.raw.data2$subject)),
    condition = levels(as.factor(reli.raw.data2$condition)),
    Trial = as.numeric(levels(as.factor(reli.raw.data2$Trial))),
    Behavior = c("Peek on top", "Peephole peek", "Right side peek", "Right box (side)")
  )

table(reli.raw.data2$IP, reli.raw.data2$subject)
```

# Interest period: toy to end of trial
```{r}
peeking_data_reli <- reli.raw.data2 %>%
  #filter(researcher_status == "Researcher absent") %>%
  filter(IP == "Toy to end of trial") %>%
  filter(Behavior %in% c("Peek on top", "Peephole peek", "Right side peek", "Right box (side)")) %>%
  mutate(duration = as.numeric(duration))%>%
  group_by(subject, condition, trial_order, Block, Trial, Behavior) %>%
  summarise(sum_duration = sum(duration), count_peeking = sum(duration>0))%>%
  ungroup()%>%
  select(-trial_order, -Block) %>%
  full_join((all_combinations_reli)) %>%
  inner_join(trialdurationdata_reli, by = c("subject", "condition", "Trial")) %>%
  complete(fill=list(sum_duration=0, count_peeking=0)) %>%
  mutate( proportion_peeking = sum_duration / Trial_duration)




table(peeking_data_reli$subject, peeking_data_reli$Trial)

write.csv(peeking_data_reli, file = "data/child_explanation_seeking_peeking_reli_data.csv")



```


### Plot of mean peeking
```{r}
peeking_data_agg_reli<-peeking_data_reli %>%
  group_by(subject, condition, Behavior) %>%
  summarise(mean.peeking = mean(sum_duration), count = sum(count_peeking))

peeking_data_agg_trial_reli<-peeking_data_reli %>%
  group_by(subject, condition, Trial, Behavior) %>%
  summarise(mean.peeking = mean(sum_duration), count = sum(count_peeking), proportion_peeking = sum(proportion_peeking))

peeking_data_agg_trial2_reli<-peeking_data_agg_trial_reli %>%
  group_by(subject, condition, Trial) %>%
  summarise(peeking_count_response = sum(count), peeking_response_binary  = sum(peeking_count_response > 0))

```

```{r}

peeking_data_proportion_reli<-peeking_data_reli %>%
  group_by(subject, condition, Trial) %>%
  summarise(prop_peeking = sum(proportion_peeking), count = sum(sum_duration > 0))
```
```{r}
reli_data<- peeking_data_agg_trial2_reli %>%
  full_join(peeking_data_proportion_reli) %>%
  select(-count)%>%
  rename(peeking_count_response_reli = peeking_count_response,
         peeking_response_binary_reli = peeking_response_binary,
         prop_peeking_reli = prop_peeking)
```
```{r}
merge_data<- peeking_data_agg_trial2 %>%
  full_join(peeking_data_proportion) %>%
  mutate(subject = recode(subject, "Es125" = "ES125")) %>%
  inner_join(reli_data)
```


# Cohen's kappa
```{r}
library(psych)

#calculate Cohen's Kappa
cohen.kappa(x=cbind(merge_data$peeking_response_binary,merge_data$peeking_response_binary_reli))


merge_data %>% summarise(match = sum(peeking_response_binary == peeking_response_binary_reli), non_match = sum(peeking_response_binary != peeking_response_binary_reli))

```

#Intraclass correlation coefficients
```{r}
merge_data2 <- as.data.frame(merge_data)
ICC(merge_data2%>%dplyr::select(prop_peeking, prop_peeking_reli), missing=TRUE)
```



