---
title: "Children Explanation Seeking Analysis Rmd"
output:
  pdf_document: default
html_document: default
date: "2023-06-13"
---
  ## Libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lme4)
library(tidyverse)
library(summarytools)
library(glmmTMB)

source("./functions/diagnostic_fcns.r")
source("./functions/glmm_stability.r")
source("./functions/boot_glmm.r")
source("./functions/boot_glmm2.r")
source("./functions/glmmTMB_stability.r")
source("./functions/drop1_para_glmmtmb.r")
source("./functions/extract_ranef_gmmTMB.r")
```


Read in the individual data files
```{r}
#Extract file names
list_csv_files <- list.files(path = "./data/children_individual_data_files/", 
               pattern = "*.xlsx", 
               full.names = T)
library("readxl")
test<-read_excel("./data/children_individual_data_files/ES_Child_101.xlsx")
orig.data <- NULL
#Read in all csv files and add the file name as variable
for(f in list_csv_files){  
  ReadInFile <- read_excel(path=f)%>%rename("file_name"=1)
  ReadInFile$child_file <- f
  orig.data <- rbind(orig.data, ReadInFile)
}

orig.data<-orig.data %>%
  rename(subject = Subject...7,
         condition = Subject...11,
         trial_order = `Trial Order`,
         behavioral_category = `Behavioral category`,
         start = `Start (s)`,
         stop= `Stop (s)`,
         duration= `Duration (s)`) %>%
  arrange(subject, start) %>%
  mutate(behavioral_category = fct_recode(as.factor(behavioral_category), "Time spent" = "Time Spent", "Researcher status" = "Researcher Status"))%>%
  mutate(Behavior = fct_recode(as.factor(Behavior), "Researcher absent" = "Researcher Absent", "Researcher present" = "Researcher Present"))%>%
  filter(!(stop=="235.969"& Behavior=="Right box (side)"& subject=="ES101"), !(duration=="4.749"& Behavior=="Right box (side)"& subject=="ES102"))

levels(orig.data$behavioral_category)
levels(as.factor(orig.data$Behavior))
```
data checks
```{r}

n.subj<-length(levels(as.factor(orig.data$subject)))
levels(as.factor(orig.data$child_file))

#add trial number
trial_data <- orig.data %>%
  filter(behavioral_category == "Trial type")

table(trial_data$subject, trial_data$FPS)

trial_data$Trial = rep(c(1:4), n.subj*2)
trial_data <- trial_data %>% select(subject, condition, behavioral_category, start, Trial)

table(trial_data$subject, trial_data$Trial)
levels(as.factor(orig.data$behavioral_category))

#add researcher status
researcher_status_data <- orig.data %>%
  filter(behavioral_category == "Researcher status")%>%
  mutate(researcher_status = Behavior) %>%
  select(subject, condition, researcher_status, behavioral_category, start)%>%
  droplevels()
table(researcher_status_data$subject, researcher_status_data$researcher_status)

#add peeking interest period 
peeking_IP_data <- orig.data %>%
  filter(Behavior == "Toy to end of trial" | Behavior == "Functional" | Behavior == "Non-functional")%>%
  mutate(IP = Behavior) %>%
  select(subject, condition, IP,start)%>%
  droplevels()

table(peeking_IP_data$subject, peeking_IP_data$IP)


orig.data2<-orig.data%>%
  full_join(trial_data)%>%
  fill(Trial) %>%
  full_join(researcher_status_data)%>%
  fill(researcher_status) %>%
  full_join(peeking_IP_data)%>%
  fill(IP)%>%
  mutate(Block = as.numeric(ifelse(
    trial_order == 1 & condition == "Functional",
    1,
    ifelse(
      trial_order == 2 & condition == "Functional",
      2,
      ifelse(
        trial_order == 1 & condition == "Non-functional",
        2,
        ifelse(trial_order == 2 &
                 condition == "Non-functional", 1, "")
      )
    )
  )))
levels(as.factor(orig.data2$Behavior))
orig.data2 %>% filter(Behavior == "Asking why") 
#happened only three times in two subjects

table(orig.data2$behavioral_category, orig.data2$Behavior)
table(orig.data2$researcher_status, orig.data2$Behavior)
table(orig.data2$subject, orig.data2$Trial)
```

## Calculate duration of each trial
```{r}
trialdurationdata <- orig.data2 %>%
  filter(Behavior == "Toy to end of trial") %>% #time chunk in which peeking should happen # old "Researcher absent"
  rename(Trial_duration = duration) %>%
  select(subject, condition, Block, Trial,  trial_order, Trial_duration)%>%
  mutate(Trial_duration = as.numeric(Trial_duration))#%>%
  # filter(!(subject=="ES117" & condition == "Non-functional" & Trial == 3))%>%#filter out two trials because the experimenter absent duration is shorter (<15 sec) or longer than expected (>40 sec)
  # filter(!(subject=="ES135" & condition == "Non-functional" & Trial == 4))

hist(trialdurationdata$Trial_duration)
mean(trialdurationdata$Trial_duration)
min(trialdurationdata$Trial_duration)

table(trialdurationdata$subject,trialdurationdata$condition)
trialdurationdata %>% group_by(condition)%>% summarise(mean_trial_duration = mean(Trial_duration), sd_trial_duration = sd(Trial_duration), se_trial_duration = sd(Trial_duration)/sqrt(length(Trial_duration)))
```

## peeking data
create dataframe just with the peeking categories and add the zeros and the trial duration

```{r}
all_combinations <-
  expand.grid(
    subject = levels(as.factor(orig.data2$subject)),
    condition = levels(as.factor(orig.data2$condition)),
    Trial = as.numeric(levels(as.factor(orig.data2$Trial))),
    Behavior = c("Peek on top", "Peephole peek", "Right side peek", "Right box (side)")
  )

table(orig.data2$IP, orig.data2$subject)
```

# Interest period: toy to end of trial
```{r}
peeking_data <- orig.data2 %>%
  #filter(researcher_status == "Researcher absent") %>%
  filter(IP == "Toy to end of trial") %>%
  filter(Behavior %in% c("Peek on top", "Peephole peek", "Right side peek", "Right box (side)")) %>%
  mutate(duration = as.numeric(duration))%>%
  group_by(subject, condition, trial_order, Block, Trial, Behavior) %>%
  summarise(sum_duration = sum(duration))%>%
  ungroup()%>%
  select(-trial_order, -Block) %>%
  full_join((all_combinations)) %>%
  inner_join(trialdurationdata, by = c("subject", "condition", "Trial")) %>%
  complete(fill=list(sum_duration=0)) %>%
  mutate(trial_overall = as.numeric(ifelse(Block == 1, Trial,
                                           ifelse(Block == 2, Trial+4, ""))),
         proportion_peeking = sum_duration / Trial_duration)




table(peeking_data$subject, peeking_data$Trial)

write.csv(peeking_data, file = "data/child_explanation_seeking_peeking_data.csv")



```


## Plot of mean peeking
```{r}
peeking_data_agg<-peeking_data %>%
  group_by(subject, condition, Behavior) %>%
  summarise(mean.peeking = mean(sum_duration), count = sum(sum_duration > 0))

peeking_data_agg_trial<-peeking_data %>%
  group_by(subject, condition, trial_overall, Behavior) %>%
  summarise(mean.peeking = mean(sum_duration), count = sum(sum_duration > 0), proportion_peeking = sum(proportion_peeking))

peeking_data_agg_trial2<-peeking_data_agg_trial %>%
  group_by(subject, condition, trial_overall) %>%
  summarise(peeking_response = sum(count > 0), peeking_response_binary  = sum(peeking_response > 0))

write.csv(peeking_data_agg, file = "Saves/Child_Overall_Peeking_Table.csv")
```


```{r}
peeking_data_agg$condition2 <- jitter(as.numeric(as.factor(peeking_data_agg$condition), amount = .0001))

#overall performance: mean peeking
ggplot(data = peeking_data_agg, aes(x = condition, y = mean.peeking)) +
  geom_boxplot(outlier.colour = "white", alpha=0.5)+ 
  geom_line(aes(x = condition2, group = subject), color = "darkgray", lty = 1, alpha = .3) +
  geom_point(aes(x = condition2))+
  theme_classic()+
  facet_wrap(~Behavior, scales = "free")
```


```{r}
#count
ggplot(data = peeking_data_agg, aes(x = condition, y = count)) +
  geom_boxplot(outlier.colour = "white")+ 
  geom_line(aes(x = condition2, group = subject), color = "darkgray", lty = 1, alpha = .3) +
  geom_point(aes(x = condition2))+
  theme_classic()+
  facet_wrap(~Behavior, scales = "free")
```


```{r}
peeking_data_agg_trial3<-peeking_data_agg_trial2 %>%
  mutate(trial_within_con = as.numeric(ifelse(trial_overall>4, trial_overall-4, trial_overall)))%>%
  group_by(condition, trial_within_con) %>%
  summarise(peeking_response_binary_summed  = sum(peeking_response_binary))
#peeking yes no
plot_binary_peeking_over_trials<-ggplot(data = peeking_data_agg_trial3, aes(x = as.factor(trial_within_con), y = peeking_response_binary_summed)) +
  geom_bar(aes(fill=condition), stat="identity", position = position_dodge(), color = "black")+
  theme_classic()+
  scale_fill_manual(values=c("Non-functional" = "darkorange", "Functional" = "dodgerblue"))+
  theme(legend.title = element_blank(), legend.position = c(0.82, 0.9))+
  labs(x = "Trial", y = "Sum of peeking individuals")+
  scale_y_continuous(breaks = seq(0, 12, 1))
plot_binary_peeking_over_trials
```

```{r}
ggsave(plot_binary_peeking_over_trials , filename = "Graphics/children_binary_peeking_count_over_trials.png", width = 6, height = 6, scale = 0.6)
```


```{r}
peeking_data_agg_trial4<-peeking_data_agg_trial2 %>%
  group_by(condition, subject) %>%
  summarise(peeking_response_binary_summed  = sum(peeking_response_binary))
```


```{r}
#peeking yes no
plot_binary_peeking_boxplot <-
  ggplot(data = peeking_data_agg_trial4, aes(x = condition, y = peeking_response_binary_summed)) +
  geom_boxplot() +
  theme_classic()

plot_binary_peeking_boxplot


peeking_data_agg_trial4$condition2 <- jitter(as.numeric(as.factor(peeking_data_agg_trial4$condition), amount = .0001))

library(gghalves)


plot_binary_peeking_boxplot <-
  ggplot(data = peeking_data_agg_trial4, aes(x = condition, y = peeking_response_binary_summed, group = condition)) +
  geom_line(
    aes(x = condition2, group = subject),
    color = "darkgray",
    lty = 1,
    alpha = .3
  ) +
  geom_point(
    data = peeking_data_agg_trial4 %>% filter(condition == "Functional"),
    aes(x = condition2),
    color = "dodgerblue",
    size = 1.5,
    alpha = .5
  ) +
  geom_point(
    data = peeking_data_agg_trial4 %>% filter(condition == "Non-functional"),
    aes(x = condition2),
    color = "darkorange",
    size = 1.5,
    alpha = .5,
   
  ) +
  geom_half_boxplot(
    data = peeking_data_agg_trial4 %>% filter(condition == "Functional"),
    aes(x = condition2, y = peeking_response_binary_summed),
    position = position_nudge(x = -0.3),
    side = "l",
    outlier.shape = NA,
    center = TRUE,
    errorbar.draw = TRUE,
    width = .05,
    fill = 'dodgerblue',
    alpha = .5
  ) +
  
  geom_half_boxplot(
    data = peeking_data_agg_trial4 %>% filter(condition == "Non-functional"),
    aes(x = condition2, y = peeking_response_binary_summed),
    position = position_nudge(x = 0.3),
    side = "r",
    outlier.shape = NA,
    center = TRUE,
    errorbar.draw = TRUE,
    width = .05,
    fill = 'darkorange',
    alpha = .5
  ) +
  
  # Define additional settings
  xlab("") +
  ylab("Sum of trial with peeking") +
  scale_x_continuous(
    breaks = c(1, 2),
    labels = c("Functional", "Non-functional"),
    limits = c(0.5, 2.5)
  ) +
  #ylim(0, 1) +
  theme_classic()

plot_binary_peeking_boxplot
```
```{r}
ggsave(plot_binary_peeking_boxplot , filename = "Graphics/children_binary_peeking_summed.png", width = 6, height = 6, scale = 0.7)

```

next steps: analyse the binary data (peeking_response_binary) with condition, Trial, and condition as predictor variables. First for all peeking combined and then separately by type (Behavior)


## Binary peeking analysis: 


### z-transformation of covariates (standardizes data)
Centering (subtract mean) and scaling (divide by SD so it becomes 1) covariates. Makes it more likely model will converge (does not change p value).
```{r}
peeking_data_agg_trial2$z.trial <-
  as.vector(scale(
    as.numeric(peeking_data_agg_trial2$trial_overall),
    scale = TRUE,
    center = TRUE
  ))
#peeking_data_agg_trial2$z.block <-
#  as.vector(scale(as.numeric(peeking_data_agg_trial2$Block)))
```

## Dummy code factors (to prepare for centering below)
makes first factor 0, second factor 1
```{r}
peeking_data_agg_trial2$condition <-
  as.factor(peeking_data_agg_trial2$condition)
peeking_data_agg_trial2$condition.dummy <-
  as.numeric(peeking_data_agg_trial2$condition == levels(peeking_data_agg_trial2$condition)[2])
```

## Center condition for random slopes (to avoid arbitrary effects of making first factor 0 and second factor 1)
```{r}
peeking_data_agg_trial2$condition.c <- as.numeric(peeking_data_agg_trial2$condition) - mean(as.numeric(peeking_data_agg_trial2$condition))
mean(peeking_data_agg_trial2$condition.c)
```

## Define control structure to make convergence more likely:
```{r}
contr <- glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000))
```

## Analysis of binary peeking response (overall peeks, includes time spent on right side)
```{r}
mm1_peeking_overall_model<-glmer(
  peeking_response_binary ~ condition + z.trial + (1 + condition.c + z.trial | subject),
  data = peeking_data_agg_trial2,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000))
)

drop1(mm1_peeking_overall_model, test = "Chisq")
summary(mm1_peeking_overall_model)
```
## zero-inflated model
```{r}
contr<-glmmTMBControl(optCtrl=list(iter.max=100000000, eval.max=100000000))

peekingmodel_binomial_zero_inflated<-glmmTMB(
  peeking_response_binary ~ condition + z.trial + (1 + condition.c + z.trial || subject),
  data = peeking_data_agg_trial2,
  ziformula=~1,
  family = binomial,
  control = contr
)
summary(peekingmodel_binomial_zero_inflated)
```

```{r}

mm1_peeking_overall_drop1 <- drop1(mm1_peeking_overall_model, test="Chisq")%>% 
  filter(!is.na(npar)) %>% 
  add_row(npar = rep(NA,1),  .before = 1) 
```

Check for colinearity
```{r}
library(car)
xx=lm(peeking_response_binary ~ condition + z.trial , data=peeking_data_agg_trial2)
vif(xx)
```

```{r}
mm1_peeking_overall_model.ci = boot.glmm.pred(
  model.res = mm1_peeking_overall_model,
  excl.warnings = F,
  nboots = 1000,
  para = T,
  n.cores = "all-1",
  resol = 1000,
  level = 0.95
)
mm1_peeking_overall_model.ci$ci.estimates
```


#### output table


```{r}
mm1_overall_peeking_output_table <-
  bind_cols(as.data.frame(summary(mm1_peeking_overall_model)$coeff),
            mm1_peeking_overall_drop1,
            #mm1_peeking_overall_model.ci$ci.estimates
            ) %>%
  select(
    Estimate,
    SE = `Std. Error`,
   # LowerCI = X2.5.,
  #  UpperCI = X97.5.,
    Chi2 = LRT,
    df = npar,
    p = `Pr(Chi)`,
    z_wald = `z value`,
    p_wald = `Pr(>|z|)`
  ) %>% #
  mutate(across(.cols = c(p), ~ format(round(.x, 3), nsmall = 3))) %>%
  mutate(across(.cols = c(Estimate:Chi2), ~ format(round(.x, 2), nsmall = 2))) %>%
  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
  mutate(p = replace(p, p == 0, "<0.001"))

write.csv(mm1_overall_peeking_output_table, file = "Saves/Child_mm1_overall_peeking_output_table.csv")
```

## First trial analysis
```{r}
peeking_data_first_trial<-peeking_data %>%
  filter(Trial == 1)%>%
  group_by(subject, condition) %>%
  summarise(sum.peeking = sum(sum_duration), count = sum(sum_duration > 0))
```

## Dummy code factors (to prepare for centering below)
makes first factor 0, second factor 1
```{r}
peeking_data_first_trial$condition <-
  as.factor(peeking_data_first_trial$condition)
peeking_data_first_trial$condition.dummy <-
  as.numeric(peeking_data_first_trial$condition == levels(peeking_data_first_trial$condition)[2])
```

## Center condition for random slopes (to avoid arbitrary effects of making first factor 0 and second factor 1)
```{r}
peeking_data_first_trial$condition.c <- as.numeric(peeking_data_first_trial$condition) - mean(as.numeric(peeking_data_first_trial$condition))
mean(peeking_data_first_trial$condition.c)
```

## Define control structure to make convergence more likely:
```{r}
contr <- glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000))
```

## Analysis of binary peeking response (overall peeks, includes time spent on right side)
```{r}
mm1_first_trial_peek_model<-glmer(
  count ~ condition + (1 + condition.c || subject),
  data = peeking_data_first_trial,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000))
)

drop1(mm1_first_trial_peek_model, test = "Chisq")
summary(mm1_first_trial_peek_model)
```


```{r}

mm1_first_trial_peek_model_drop1 <- drop1(mm1_first_trial_peek_model, test="Chisq")%>% 
  filter(!is.na(npar)) %>% 
  add_row(npar = rep(NA,1),  .before = 1) 
```

```{r}
mm1_first_trial_peek_model.ci = boot.glmm.pred(
  model.res = mm1_first_trial_peek_model,
  excl.warnings = F,
  nboots = 1000,
  para = T,
  n.cores = "all-1",
  resol = 1000,
  level = 0.95
)
mm1_first_trial_peek_model.ci$ci.estimates
```


#### output table


```{r}
mm1_first_trial_peek_model_output_table <-
  bind_cols(as.data.frame(summary(mm1_first_trial_peek_model)$coeff),
            mm1_first_trial_peek_model_drop1,
            #mm1_first_trial_peek_model.ci$ci.estimates
            ) %>%
  select(
    Estimate,
    SE = `Std. Error`,
   # LowerCI = X2.5.,
  #  UpperCI = X97.5.,
    Chi2 = LRT,
    df = npar,
    p = `Pr(Chi)`,
    z_wald = `z value`,
    p_wald = `Pr(>|z|)`
  ) %>% #
  mutate(across(.cols = c(p), ~ format(round(.x, 3), nsmall = 3))) %>%
  mutate(across(.cols = c(Estimate:Chi2), ~ format(round(.x, 2), nsmall = 2))) %>%
  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
  mutate(p = replace(p, p == 0, "<0.001"))

write.csv(mm1_first_trial_peek_model_output_table, file = "Saves/Child_mm1_first_trial_peeking_output_table.csv")
```

## Proportion analysis

```{r}
library(glmmTMB)
library(car)

source("./functions/diagnostic_fcns.r")
source("./functions/glmm_stability.r")
source("./functions/boot_glmmTMB.r")
source("./functions/glmmTMB_stability.r")
source("./functions/drop1_para_glmmtmb.r")
source("./functions/extract_ranef_gmmTMB.r")
```

```{r}
peeking_data_proportion<-peeking_data %>%
  group_by(subject, condition, trial_overall) %>%
  summarise(prop_peeking = sum(proportion_peeking), count = sum(sum_duration > 0))
```
```{r}
library(ggbeeswarm)

plot_peeking_data_proportion_over_trials <-
  ggplot(data = peeking_data_proportion, aes(
    x = as.factor(trial_overall),
    y = prop_peeking,
    color = condition
  )) +
  geom_boxplot(aes(), outlier.colour = "white") +
  geom_quasirandom(dodge.width = 0.8, alpha = 0.3) +
  theme_classic() +
  scale_color_manual(values = c(
    "Non-functional" = "darkorange",
    "Functional" = "dodgerblue"
  )) +
  theme(legend.title = element_blank(),
        legend.position = c(0.1, 0.9)) +
  labs(x = "Trial", y = "Proportion time peeking")
plot_peeking_data_proportion_over_trials
```

```{r}
ggsave(plot_binary_peeking_over_trials , filename = "Graphics/children_binary_peeking_count_over_trials.png", width = 8, height = 6, scale = 0.6)
```

### create plot grid
```{r}
library(cowplot)
pg_proportion<- plot_grid(plot_peeking_data_proportion_over_trials, plot_peeking_data_proportion_over_trials, labels = c("A", "B"), nrow = 1)

pg_proportion
```

### Transformation check of ALL PEEKS proportion 
```{r}
contr<-glmmTMBControl(optCtrl=list(iter.max=100000000, eval.max=100000000))
peeking_data_proportion$prop_peeking_scaled <- (peeking_data_proportion$prop_peeking*(length(peeking_data_proportion$prop_peeking) - 1) + 0.5)/length(peeking_data_proportion$prop_peeking)
```

### z-transformation of covariates (standardizes data)
Centering (subtract mean) and scaling (divide by SD so it becomes 1) covariates. Makes it more likely model will converge (does not change p value).
```{r}
peeking_data_proportion$z.trial <-
  as.vector(scale(
    as.numeric(peeking_data_agg_trial2$trial_overall),
    scale = TRUE,
    center = TRUE
  ))
```

## Dummy code factors (to prepare for centering below)
makes first factor 0, second factor 1
```{r}
peeking_data_proportion$condition <-
  as.factor(peeking_data_proportion$condition)
peeking_data_proportion$condition.dummy <-
  as.numeric(peeking_data_proportion$condition == levels(peeking_data_proportion$condition)[2])
```

## Center condition for random slopes (to avoid arbitrary effects of making first factor 0 and second factor 1)
```{r}
peeking_data_proportion$condition.c <- as.numeric(peeking_data_proportion$condition) - mean(as.numeric(peeking_data_proportion$condition))
mean(peeking_data_proportion$condition.c)
```

### Analysis of proportional ALL peeking response
```{r}
peekingmodel_beta<-glmmTMB(
  prop_peeking_scaled ~ condition + z.trial + 
    (1 + condition.c + z.trial || subject), #remove correlation between random slope and intercept due to convergence warning
  data = peeking_data_proportion,
  #ziformula=~1,
  family = beta_family,
  control = contr
)
```

## check if model is overdispersed or not (assumption check for proportion models)
```{r}
overdisp.test(peekingmodel_beta)
```


```{r}
summary(peekingmodel_beta)

drop1(peekingmodel_beta, test="Chisq")
```
LRT
```{r}
peekingmodel_beta_drop1 <- drop1(peekingmodel_beta, test="Chisq")%>% 
  filter(!is.na(Df)) %>% 
  add_row(Df = rep(NA,1),  .before = 1) 
```


+ Collinearity checks
```{r}
library(car)
xx=lm(prop_peeking_scaled ~ condition + z.trial, data=peeking_data_proportion)
vif(xx)
```
confidence intervals
```{r eval=FALSE}
peekingmodel_beta.ci=boot.glmmtmb(peekingmodel_beta, 
                                  nboots=1000, para=T, n.cores="all-1", resol=1000, level=0.95, data=peeking_data_proportion)
```

#### output table

```{r}
peekingmodel_beta_table <- bind_cols(as.data.frame(summary(peekingmodel_beta)$coefficients$cond),
                                     peekingmodel_beta_drop1)%>%#,
  # mm1_area_tail.ci$ci.estimates$fe[1:7,]) %>%
  dplyr::select(Estimate, SE = `Std. Error`, Chi2 = LRT, df = Df, p = `Pr(>Chi)`, z_wald=`z value`, p_wald=`Pr(>|z|)`) %>% #LowerCI = X2.5., UpperCI = X97.5., 
  mutate(across(.cols = c(p), ~ format(round(.x, 3), nsmall=3))) %>% 
  mutate(across(.cols = c(Estimate:Chi2), ~ format(round(.x, 2), nsmall=2))) %>% 
  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
  mutate(p=replace(p, p==0, "<0.001"))

write.csv(peekingmodel_beta_table , file = "saves/Children_peekingmodel_beta_table.csv")
```



# Interest period: Researcher absent
```{r}
peeking_data_researcher_absent <- orig.data2 %>%
  filter(researcher_status == "Researcher absent") %>%
  filter(Behavior %in% c("Peek on top", "Peephole peek", "Right side peek", "Right box (side)")) %>%
  mutate(duration = as.numeric(duration))%>%
  group_by(subject, condition, trial_order, Block, Trial, Behavior) %>%
  summarise(sum_duration = sum(duration))%>%
  ungroup()%>%
  select(-trial_order, -Block) %>%
  full_join((all_combinations)) %>%
  inner_join(trialdurationdata, by = c("subject", "condition", "Trial")) %>%
  complete(fill=list(sum_duration=0)) %>%
  mutate(trial_overall = as.numeric(ifelse(Block == 1, Trial,
                                           ifelse(Block == 2, Trial+4, ""))),
         proportion_peeking = sum_duration / Trial_duration)


write.csv(peeking_data_researcher_absent, file = "data/child_explanation_seeking_peeking_data_researcher_absent.csv")



```


## Plot of mean peeking
```{r}
peeking_data_agg_researcher_absent<-peeking_data_researcher_absent %>%
  group_by(subject, condition, Behavior) %>%
  summarise(mean.peeking = mean(sum_duration), count = sum(sum_duration > 0))

peeking_data_agg_trial_researcher_absent<-peeking_data_researcher_absent %>%
  group_by(subject, condition, trial_overall, Behavior) %>%
  summarise(mean.peeking = mean(sum_duration), count = sum(sum_duration > 0), proportion_peeking = sum(proportion_peeking))

peeking_data_agg_trial2_researcher_absent<-peeking_data_agg_trial_researcher_absent %>%
  group_by(subject, condition, trial_overall) %>%
  summarise(peeking_response = sum(count > 0), peeking_response_binary  = sum(peeking_response > 0))

write.csv(peeking_data_agg_researcher_absent, file = "Saves/Child_Overall_Peeking_Table.csv")
```


```{r}
peeking_data_agg_researcher_absent$condition2 <- jitter(as.numeric(as.factor(peeking_data_agg_researcher_absent$condition), amount = .0001))

#overall performance: mean peeking
ggplot(data = peeking_data_agg_researcher_absent, aes(x = condition, y = mean.peeking)) +
  geom_boxplot(outlier.colour = "white", alpha=0.5)+ 
  geom_line(aes(x = condition2, group = subject), color = "darkgray", lty = 1, alpha = .3) +
  geom_point(aes(x = condition2))+
  theme_classic()+
  facet_wrap(~Behavior, scales = "free")
```


```{r}
#count
ggplot(data = peeking_data_agg_researcher_absent, aes(x = condition, y = count)) +
  geom_boxplot(outlier.colour = "white")+ 
  geom_line(aes(x = condition2, group = subject), color = "darkgray", lty = 1, alpha = .3) +
  geom_point(aes(x = condition2))+
  theme_classic()+
  facet_wrap(~Behavior, scales = "free")
```


```{r}
peeking_data_agg_trial3_researcher_absent<-peeking_data_agg_trial2_researcher_absent %>%
  group_by(condition, trial_overall) %>%
  summarise(peeking_response_binary_summed  = sum(peeking_response_binary))
#peeking yes no
plot_binary_peeking_over_trials<-ggplot(data = peeking_data_agg_trial3_researcher_absent, aes(x = trial_overall, y = peeking_response_binary_summed)) +
  geom_bar(aes(fill=condition), stat="identity", position = position_dodge())+
  theme_classic()

plot_binary_peeking_over_trials
```
```{r}
ggsave(plot_binary_peeking_over_trials, filename = "Graphics/children_binary_peeking_count_over_trials_researcher_absent.png", width = 8, height = 6, scale = 0.7)
```

## Binary peeking analysis: 


### z-transformation of covariates (standardizes data)
Centering (subtract mean) and scaling (divide by SD so it becomes 1) covariates. Makes it more likely model will converge (does not change p value).
```{r}
peeking_data_agg_trial2_researcher_absent$z.trial <-
  as.vector(scale(
    as.numeric(peeking_data_agg_trial2$trial_overall),
    scale = TRUE,
    center = TRUE
  ))
#peeking_data_agg_trial2$z.block <-
#  as.vector(scale(as.numeric(peeking_data_agg_trial2$Block)))
```

## Dummy code factors (to prepare for centering below)
makes first factor 0, second factor 1
```{r}
peeking_data_agg_trial2_researcher_absent$condition <-
  as.factor(peeking_data_agg_trial2_researcher_absent$condition)
peeking_data_agg_trial2_researcher_absent$condition.dummy <-
  as.numeric(peeking_data_agg_trial2_researcher_absent$condition == levels(peeking_data_agg_trial2_researcher_absent$condition)[2])
```

## Center condition for random slopes (to avoid arbitrary effects of making first factor 0 and second factor 1)
```{r}
peeking_data_agg_trial2_researcher_absent$condition.c <- as.numeric(peeking_data_agg_trial2_researcher_absent$condition) - mean(as.numeric(peeking_data_agg_trial2_researcher_absent$condition))
mean(peeking_data_agg_trial2_researcher_absent$condition.c)
```

## Define control structure to make convergence more likely:
```{r}
contr <- glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000))
```

## Analysis of binary peeking response (overall peeks, includes time spent on right side)
```{r}
mm1_peeking_overall_model_researcher_absent<-glmer(
  peeking_response_binary ~ condition + z.trial + (1 + condition.c + z.trial | subject),
  data = peeking_data_agg_trial2_researcher_absent,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000))
)

drop1(mm1_peeking_overall_model_researcher_absent, test = "Chisq")
summary(mm1_peeking_overall_model_researcher_absent)
```


```{r}

mm1_peeking_overall_researcher_absent_drop1 <- drop1(mm1_peeking_overall_model_researcher_absent, test="Chisq")%>% 
  filter(!is.na(npar)) %>% 
  add_row(npar = rep(NA,1),  .before = 1) 
```

Check for colinearity
```{r}
library(car)
xx=lm(peeking_response_binary ~ condition + z.trial , data=peeking_data_agg_trial2_researcher_absent)
vif(xx)
```

```{r}
mm1_peeking_overall_model_researcher_absent.ci = boot.glmm.pred(
  model.res = mm1_peeking_overall_model_researcher_absent,
  excl.warnings = F,
  nboots = 1000,
  para = T,
  n.cores = "all-1",
  resol = 1000,
  level = 0.95
)
mm1_peeking_overall_model_researcher_absent.ci$ci.estimates
```


#### output table


```{r}
mm1_overall_peeking_researcher_absent_output_table <-
  bind_cols(as.data.frame(summary(mm1_peeking_overall_model_researcher_absent)$coeff),
            mm1_peeking_overall_researcher_absent_drop1,
            #mm1_peeking_overall_model.ci$ci.estimates
            ) %>%
  select(
    Estimate,
    SE = `Std. Error`,
   # LowerCI = X2.5.,
  #  UpperCI = X97.5.,
    Chi2 = LRT,
    df = npar,
    p = `Pr(Chi)`,
    z_wald = `z value`,
    p_wald = `Pr(>|z|)`
  ) %>% #
  mutate(across(.cols = c(p), ~ format(round(.x, 3), nsmall = 3))) %>%
  mutate(across(.cols = c(Estimate:Chi2), ~ format(round(.x, 2), nsmall = 2))) %>%
  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
  mutate(p = replace(p, p == 0, "<0.001"))

write.csv(mm1_overall_peeking_researcher_absent_output_table, file = "Saves/Child_mm1_overall_peeking_researcher_absent_output_table.csv")
```

## Proportion analysis

```{r}
library(glmmTMB)
library(car)

source("./functions/diagnostic_fcns.r")
source("./functions/glmm_stability.r")
source("./functions/boot_glmmTMB.r")
source("./functions/glmmTMB_stability.r")
source("./functions/drop1_para_glmmtmb.r")
source("./functions/extract_ranef_gmmTMB.r")
```
```{r}
peeking_data_proportion_researcher_absent<-peeking_data_researcher_absent %>%
  group_by(subject, condition, trial_overall) %>%
  summarise(prop_peeking = sum(proportion_peeking), count = sum(sum_duration > 0))
```

### Transformation check of ALL PEEKS proportion 
```{r}
contr<-glmmTMBControl(optCtrl=list(iter.max=100000000, eval.max=100000000))
peeking_data_proportion_researcher_absent$prop_peeking_scaled <- (peeking_data_proportion_researcher_absent$prop_peeking*(length(peeking_data_proportion_researcher_absent$prop_peeking) - 1) + 0.5)/length(peeking_data_proportion_researcher_absent$prop_peeking)
```

### z-transformation of covariates (standardizes data)
Centering (subtract mean) and scaling (divide by SD so it becomes 1) covariates. Makes it more likely model will converge (does not change p value).
```{r}
peeking_data_proportion_researcher_absent$z.trial <-
  as.vector(scale(
    as.numeric(peeking_data_agg_trial2_researcher_absent$trial_overall),
    scale = TRUE,
    center = TRUE
  ))
```

## Dummy code factors (to prepare for centering below)
makes first factor 0, second factor 1
```{r}
peeking_data_proportion_researcher_absent$condition <-
  as.factor(peeking_data_proportion_researcher_absent$condition)
peeking_data_proportion_researcher_absent$condition.dummy <-
  as.numeric(peeking_data_proportion_researcher_absent$condition == levels(peeking_data_proportion_researcher_absent$condition)[2])
```

## Center condition for random slopes (to avoid arbitrary effects of making first factor 0 and second factor 1)
```{r}
peeking_data_proportion_researcher_absent$condition.c <- as.numeric(peeking_data_proportion_researcher_absent$condition) - mean(as.numeric(peeking_data_proportion_researcher_absent$condition))
mean(peeking_data_proportion_researcher_absent$condition.c)
```

### Analysis of proportional ALL peeking response
```{r}
peekingmodel_beta_researcher_absent<-glmmTMB(
  prop_peeking_scaled ~ condition + z.trial + 
    (1 + condition.c + z.trial || subject), #remove correlation between random slope and intercept due to convergence warning
  data = peeking_data_proportion_researcher_absent,
  #ziformula=~1,
  family = beta_family,
  control = contr
)
```

## check if model is overdispersed or not (assumption check for proportion models)
```{r}
overdisp.test(peekingmodel_beta_researcher_absent)
```


```{r}
summary(peekingmodel_beta_researcher_absent)

drop1(peekingmodel_beta_researcher_absent, test="Chisq")
```
LRT
```{r}
peekingmodel_beta_researcher_absent_drop1 <- drop1(peekingmodel_beta_researcher_absent, test="Chisq")%>% 
  filter(!is.na(Df)) %>% 
  add_row(Df = rep(NA,1),  .before = 1) 
```


+ Collinearity checks
```{r}
library(car)
xx=lm(prop_peeking_scaled ~ condition + z.trial, data=peeking_data_proportion_researcher_absent)
vif(xx)
```
confidence intervals
```{r eval=FALSE}
peekingmodel_beta_researcher_absent.ci=boot.glmmtmb(peekingmodel_beta_researcher_absent, 
                                  nboots=1000, para=T, n.cores="all-1", resol=1000, level=0.95, data=peeking_data_proportion_researcher_absent)
```

#### output table

```{r}
peekingmodel_beta_researcher_absent_table <- bind_cols(as.data.frame(summary(peekingmodel_beta_researcher_absent)$coefficients$cond),
                                     peekingmodel_beta_researcher_absent_drop1)%>%#,
  # mm1_area_tail.ci$ci.estimates$fe[1:7,]) %>%
  dplyr::select(Estimate, SE = `Std. Error`, Chi2 = LRT, df = Df, p = `Pr(>Chi)`, z_wald=`z value`, p_wald=`Pr(>|z|)`) %>% #LowerCI = X2.5., UpperCI = X97.5., 
  mutate(across(.cols = c(p), ~ format(round(.x, 3), nsmall=3))) %>% 
  mutate(across(.cols = c(Estimate:Chi2), ~ format(round(.x, 2), nsmall=2))) %>% 
  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
  mutate(p=replace(p, p==0, "<0.001"))

write.csv(peekingmodel_beta_researcher_absent_table , file = "saves/Children_peekingmodel_beta_researcher_absent_table.csv")
``` 






## Plotting

### toy to end of trial
```{r}
#peeking_data_agg_trial2.agg$condition<-fct_relevel(as.factor(peeking_data_agg_trial2.agg$condition), "blocked", "teasing", "clumsy")

peeking_data_agg_trial2.plot <- peeking_data_agg_trial2 %>%
  group_by(subject, condition)%>%
  summarise(mean_binary_peek = mean(peeking_response_binary), peeking_response_binary_summed  = sum(peeking_response_binary))

peeking_data_agg_trial2.plot$condition2 <- jitter(as.numeric(as.factor(peeking_data_agg_trial2.plot$condition), amount = .0001))

library(gghalves)


sum_peeking_plot <-
  ggplot(data = peeking_data_agg_trial2.plot, aes(x = condition, y = peeking_response_binary_summed, group = condition)) +
  geom_line(
    aes(x = condition2, group = subject),
    color = "darkgray",
    lty = 1,
    alpha = .3
  ) +
  geom_point(
    data = peeking_data_agg_trial2.plot %>% filter(condition == "Functional"),
    aes(x = condition2),
    color = "dodgerblue",
    size = 1.5,
    alpha = .5
  ) +
  geom_point(
    data = peeking_data_agg_trial2.plot %>% filter(condition == "Non-functional"),
    aes(x = condition2),
    color = "darkorange",
    size = 1.5,
    alpha = .5,
    
  ) +
  geom_half_boxplot(
    data = peeking_data_agg_trial2.plot %>% filter(condition == "Functional"),
    aes(x = condition2, y = mean_binary_peek),
    position = position_nudge(x = -0.3),
    side = "l",
    outlier.shape = NA,
    center = TRUE,
    errorbar.draw = TRUE,
    width = .05,
    fill = 'dodgerblue',
    alpha = .5
  ) +
  
  geom_half_boxplot(
    data = peeking_data_agg_trial2.plot %>% filter(condition == "Non-functional"),
    aes(x = condition2, y = mean_binary_peek),
    position = position_nudge(x = 0.3),
    side = "r",
    outlier.shape = NA,
    center = TRUE,
    errorbar.draw = TRUE,
    width = .05,
    fill = 'darkorange',
    alpha = .5
  ) +
  
  # Define additional settings
  xlab("") +
  ylab("Proportion of trial with peeking") +
  scale_x_continuous(
    breaks = c(1, 2),
    labels = c("Functional", "Non-functional"),
    limits = c(0.5, 2.5)
  ) +
  #ylim(0, 1) +
  theme_classic()

sum_peeking_plot
```


```{r}
ggsave(sum_peeking_plot , filename = "Graphics/children_binary_peeking_plot_toy_to_end_of_trial.png", width = 6, height = 6, scale = 0.7)

```



### researchers absent IP
```{r}
#peeking_data_agg_trial2.agg$condition<-fct_relevel(as.factor(peeking_data_agg_trial2.agg$condition), "blocked", "teasing", "clumsy")

peeking_data_agg_trial2_researcher_absent.plot <- peeking_data_agg_trial2_researcher_absent %>%
  group_by(subject, condition)%>%
  summarise(mean_binary_peek = mean(peeking_response_binary), peeking_response_binary_summed  = sum(peeking_response_binary))

peeking_data_agg_trial2_researcher_absent.plot$condition2 <- jitter(as.numeric(as.factor(peeking_data_agg_trial2_researcher_absent.plot$condition), amount = .0001))

library(gghalves)


child_sum_peeking_plot_researcher_absent <-
  ggplot(data = peeking_data_agg_trial2_researcher_absent.plot, aes(x = condition, y = peeking_response_binary_summed, group = condition)) +
  geom_line(
    aes(x = condition2, group = subject),
    color = "darkgray",
    lty = 1,
    alpha = .3
  ) +
  geom_point(
    data = peeking_data_agg_trial2_researcher_absent.plot %>% filter(condition == "Functional"),
    aes(x = condition2),
    color = "dodgerblue",
    size = 1.5,
    alpha = .5
  ) +
  geom_point(
    data = peeking_data_agg_trial2_researcher_absent.plot %>% filter(condition == "Non-functional"),
    aes(x = condition2),
    color = "darkorange",
    size = 1.5,
    alpha = .5,
    
  ) +
  geom_half_boxplot(
    data = peeking_data_agg_trial2_researcher_absent.plot %>% filter(condition == "Functional"),
    aes(x = condition2, y = peeking_response_binary_summed),
    position = position_nudge(x = -0.3),
    side = "l",
    outlier.shape = NA,
    center = TRUE,
    errorbar.draw = TRUE,
    width = .05,
    fill = 'dodgerblue',
    alpha = .5
  ) +
  
  geom_half_boxplot(
    data = peeking_data_agg_trial2_researcher_absent.plot %>% filter(condition == "Non-functional"),
    aes(x = condition2, y = peeking_response_binary_summed),
    position = position_nudge(x = 0.3),
    side = "r",
    outlier.shape = NA,
    center = TRUE,
    errorbar.draw = TRUE,
    width = .05,
    fill = 'darkorange',
    alpha = .5
  ) +
  
  # Define additional settings
  xlab("") +
  ylab("Sum of trials with peeking") +
  scale_x_continuous(
    breaks = c(1, 2),
    labels = c("Functional", "Non-functional"),
    limits = c(0.5, 2.5)
  ) +
  #ylim(0, 1) +
  theme_classic()

child_sum_peeking_plot_researcher_absent
```
```{r}
ggsave(sum_peeking_plot_researcher_absent , filename = "Graphics/children_binary_peeking_plot_researcher_absent.png", width = 6, height = 6, scale = 0.7)

```


```{r}
library(cowplot)
plot_grid(sum_peeking_plot, sum_peeking_plot_researcher_absent)
```


### Plot: Mean peeking duration
```{r}
peeking_data.plot <- peeking_data %>%
  group_by(subject, condition) %>%
  summarise(mean_duration_peeking = mean(sum_duration, na.rm=TRUE),
            mean_prop_duration_peeking = mean(proportion_peeking, na.rm=TRUE))
```


```{r}
#xdata.agg$condition<-fct_relevel(as.factor(xdata.agg$condition), "blocked", "teasing", "clumsy")
peeking_data.plot$condition2 <- jitter(as.numeric(as.factor(peeking_data.plot$condition), amount = .0001))

library(gghalves)


sum_peeking_plot <-
  ggplot(data = peeking_data.plot, aes(x = condition, y = mean_duration_peeking, group = condition)) +
  geom_line(
    aes(x = condition2, group = subject),
    color = "darkgray",
    lty = 1,
    alpha = .3
  ) +
  geom_point(
    data = peeking_data.plot %>% filter(condition == "Functional"),
    aes(x = condition2),
    color = "dodgerblue",
    size = 1.5,
    alpha = .5
  ) +
  geom_point(
    data = peeking_data.plot %>% filter(condition == "Non-functional"),
    aes(x = condition2),
    color = "darkorange",
    size = 1.5,
    alpha = .5,
    
  ) +
  geom_half_boxplot(
    data = peeking_data.plot %>% filter(condition == "Functional"),
    aes(x = condition2, y = mean_duration_peeking),
    position = position_nudge(x = -0.3),
    side = "l",
    outlier.shape = NA,
    center = TRUE,
    errorbar.draw = TRUE,
    width = .05,
    fill = 'dodgerblue',
    alpha = .5
  ) +
  
  geom_half_boxplot(
    data = peeking_data.plot %>% filter(condition == "Non-functional"),
    aes(x = condition2, y = mean_duration_peeking),
    position = position_nudge(x = 0.3),
    side = "r",
    outlier.shape = NA,
    center = TRUE,
    errorbar.draw = TRUE,
    width = .05,
    fill = 'darkorange',
    alpha = .5
  ) +
  
  # Define additional settings
  xlab("") +
  ylab("Mean peeking duration (s)") +
  scale_x_continuous(
    breaks = c(1, 2),
    labels = c("Functional", "Non-functional"),
    limits = c(0.5, 2.5)
  ) +
  #ylim(0, 1) +
  theme_classic()

sum_peeking_plot 

ggsave(sum_peeking_plot , filename = "Graphics/children_sum_peeking_plot.png", width = 6, height = 6, scale = 0.6)

```
### Plot: Mean proportion peeking duration
```{r}
#xdata.agg$condition<-fct_relevel(as.factor(xdata.agg$condition), "blocked", "teasing", "clumsy")
peeking_data.plot$condition2 <- jitter(as.numeric(as.factor(peeking_data.plot$condition), amount = .0001))

library(gghalves)


sum_prop_peeking_plot <-
  ggplot(data = peeking_data.plot, aes(x = condition, y = mean_prop_duration_peeking, group = condition)) +
  geom_line(
    aes(x = condition2, group = subject),
    color = "darkgray",
    lty = 1,
    alpha = .3
  ) +
  geom_point(
    data = peeking_data.plot %>% filter(condition == "Functional"),
    aes(x = condition2),
    color = "dodgerblue",
    size = 1.5,
    alpha = .5
  ) +
  geom_point(
    data = peeking_data.plot %>% filter(condition == "Non-functional"),
    aes(x = condition2),
    color = "darkorange",
    size = 1.5,
    alpha = .5,
    
  ) +
  geom_half_boxplot(
    data = peeking_data.plot %>% filter(condition == "Functional"),
    aes(x = condition2, y = mean_prop_duration_peeking),
    position = position_nudge(x = -0.3),
    side = "l",
    outlier.shape = NA,
    center = TRUE,
    errorbar.draw = TRUE,
    width = .05,
    fill = 'dodgerblue',
    alpha = .5
  ) +
  
  geom_half_boxplot(
    data = peeking_data.plot %>% filter(condition == "Non-functional"),
    aes(x = condition2, y = mean_prop_duration_peeking),
    position = position_nudge(x = 0.3),
    side = "r",
    outlier.shape = NA,
    center = TRUE,
    errorbar.draw = TRUE,
    width = .05,
    fill = 'darkorange',
    alpha = .5
  ) +
  
  # Define additional settings
  xlab("") +
  ylab("Mean proportion peeking duration") +
  scale_x_continuous(
    breaks = c(1, 2),
    labels = c("Functional", "Non-functional"),
    limits = c(0.5, 2.5)
  ) +
  #ylim(0, 1) +
  theme_classic()

sum_prop_peeking_plot 

ggsave(sum_prop_peeking_plot , filename = "Graphics/children_sum_prop_peeking_plot.png", width = 6, height = 6, scale = 0.6)

```












