---
title: "Kea Explanation Seeking Analysis Rmd"
output:
  pdf_document: default
  html_document: default
date: "2023-02-07"
---
## Libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lme4)
library(tidyverse)
library(summarytools)


source("./functions/diagnostic_fcns.r")
source("./functions/glmm_stability.r")
source("./functions/boot_glmm.r")
source("./functions/boot_glmm2.r")
source("./functions/glmmTMB_stability.r")
source("./functions/drop1_para_glmmtmb.r")
source("./functions/extract_ranef_gmmTMB.r")
```

## Read in data
```{r}
raw.data<-read.csv("data/ALL SUBJECTS Explanation Seeking Coding 17.1.2023 csv.csv")
cb.data<-read.csv("data/counter balancing trials.csv")%>%
  select(-Date)%>%
  filter(Condition != "Refresher")
cb.data2<-expand.grid(Subject=levels(as.factor(cb.data$Subject)), Trial=c(4,5), Session=c(1:8), Behavior=c("Right side peek", "Peek on top", "Peephole peek"))
```

## Calculate duration of each trial
```{r}
trialdurationdata<-raw.data %>%
  filter(Behavior == "Food to end of trial") %>%
  rename(Condition = Subject.1, Trial_Duration = Duration..s.) %>%
  mutate(Trial = as.numeric(ifelse(Trial.Order == 1 & Condition == "Functional", 4, 
                                   ifelse(Trial.Order == 2 & Condition == "Functional", 5,
                                          ifelse(Trial.Order == 1 & Condition == "Non-functional", 5,
                                                 ifelse(Trial.Order == 2 & Condition == "Non-functional", 4,""))))))%>%
  select(Subject, Condition, Session, Trial,  Trial_Duration)

table(trialdurationdata$Subject,trialdurationdata$Condition)
table(cb.data$Subject,cb.data$Condition)
```
## Subset the raw data to the exploration period
first we determine when the food hit the ground and then we filter the data such that we only keep the peeks following the food release
```{r}
solution.data<-raw.data %>%
  filter( Behavior =="Food to end of trial") %>%
  rename(Condition = Subject.1) %>%
  mutate(Trial = as.numeric(ifelse(Trial.Order == 1 & Condition == "Functional", 4, 
                        ifelse(Trial.Order == 2 & Condition == "Functional", 5,
                               ifelse(Trial.Order == 1 & Condition == "Non-functional", 5,
                                      ifelse(Trial.Order == 2 & Condition == "Non-functional", 4,""))))))%>%
  group_by(Subject, Session, Trial, Condition)%>% #Subject.1 means trial type (func or non func)
  summarise(solution_time=min(Start..s.))


raw.data.after.solution <- raw.data %>%
  rename(Condition = Subject.1) %>%
  mutate(Trial = as.numeric(ifelse(Trial.Order == 1 & Condition == "Functional", 4, 
                        ifelse(Trial.Order == 2 & Condition == "Functional", 5,
                               ifelse(Trial.Order == 1 & Condition == "Non-functional", 5,
                                      ifelse(Trial.Order == 2 & Condition == "Non-functional", 4,""))))))%>%
  full_join(solution.data) %>%
  filter(Start..s.>=solution_time)
```


## Data preparation (summed up peeks, trial duration, added 0s, proportion peeking)
```{r}
xdata<-raw.data.after.solution %>%
  filter(Behavioral.category == "Peeking" | (Behavioral.category == "Proximity" & Behavior == "Right box (side)")) %>%
  group_by(Subject, Session, Trial, Condition)%>% #Subject.1 means trial type (func or non func)
  summarise(sum_peeking=sum(Duration..s.))%>% #aggregating durations per behavioral category within each session
  ungroup()%>%
  full_join(cb.data)%>%
  complete(Subject, Session, Trial,  fill=list(sum_peeking=0))%>% #fill in 0s
  full_join(trialdurationdata)%>%
  mutate(Proportionpeeking = sum_peeking / Trial_Duration,
         peeking_binary= as.numeric(ifelse(sum_peeking > 0, 1, 0)))
```

## Data frame (types of peeking split up)
```{r}
peekingdata<-raw.data %>%
  filter(Behavioral.category == "Peeking" | (Behavioral.category == "Proximity" & Behavior == "Right box (side)")) %>%
rename(Condition = Subject.1) %>%
  mutate(Trial = as.numeric(ifelse(Trial.Order == 1 & Condition == "Functional", 4,
                                   ifelse(Trial.Order == 2 & Condition == "Functional", 5,
                                          ifelse(Trial.Order == 1 & Condition == "Non-functional", 5,
                                                 ifelse(Trial.Order == 2 & Condition == "Non-functional", 4,""))))))%>%
  group_by(Subject, Session, Trial, Condition, Behavior)%>%
  summarise(sum_peeking=sum(Duration..s.))%>% #aggregating durations per behavioral category within each session
  ungroup()%>%
  full_join(cb.data2)%>%
  complete(Subject, Session, Trial, Behavior,  fill=list(sum_peeking=0))%>% #fill in 0s
  select(-Condition)%>%
  full_join(cb.data)%>%
  full_join(trialdurationdata)%>%
  mutate(Proportionpeeking = sum_peeking / Trial_Duration,
         peeking_binary= as.numeric(ifelse(sum_peeking > 0, 1, 0)))

view(dfSummary(peekingdata)) #data frame summary of peeking data

xdata%>%filter(is.na(Proportionpeeking))
view(dfSummary(xdata)) #data frame summary of xdata
```

## Plot of mean peeking
```{r}
xdata.agg2<-xdata %>%
  group_by(Subject, Condition) %>%
  summarise(mean.peeking = mean(Proportionpeeking), count = sum(Proportionpeeking > 0))

write.csv(xdata.agg2, file = "Saves/Overall Peeking Table.csv")

xdata.agg<-xdata %>%
  group_by(Subject, Condition) %>%
  summarise(mean.peeking = mean(sum_peeking), count = sum(Proportionpeeking > 0), mean.prop.peeking = mean(Proportionpeeking))

xdata.agg%>%
  group_by(Condition)%>%
  summarise(sum(count>0), sum(count==0))

xdata.agg$condition2 <- jitter(as.numeric(as.factor(xdata.agg$Condition), amount = .0001))

#overall performance
ggplot(data = xdata.agg, aes(x = Condition, y = mean.peeking)) +
  geom_boxplot(outlier.colour = "white")+ 
  geom_line(aes(x = condition2, group = Subject), color = "darkgray", lty = 1, alpha = .3) +
  geom_point(aes(x = condition2))+
  theme_classic()#+


ggplot(data = xdata.agg, aes(x = Condition, y = count)) +
  geom_boxplot(outlier.colour = "white")+ 
  geom_line(aes(x = condition2, group = Subject), color = "darkgray", lty = 1, alpha = .3) +
  geom_point(aes(x = condition2))+
  theme_classic()+
  facet_wrap(~Subject)


#tool users

ggplot(data = xdata.agg%>%filter(Subject=="Paul" | Subject=="John" | Subject=="Kermit" | Subject=="Frowin" | Subject =="Pick"), aes(x = Condition, y = mean.peeking)) +
  geom_boxplot(outlier.colour = "white")+ 
  geom_line(aes(x = condition2, group = Subject), color = "darkgray", lty = 1, alpha = .3) +
  geom_point(aes(x = condition2))+
  theme_classic()#+



```
## Mean peeking duration
```{r}
#xdata.agg$condition<-fct_relevel(as.factor(xdata.agg$condition), "blocked", "teasing", "clumsy")
xdata.agg$condition2 <- jitter(as.numeric(as.factor(xdata.agg$Condition), amount = .0001))

library(gghalves)


sum_peeking_plot <-
  ggplot(data = xdata.agg, aes(x = Condition, y = mean.peeking, group = Condition)) +
  geom_line(
    aes(x = condition2, group = Subject),
    color = "darkgray",
    lty = 1,
    alpha = .3
  ) +
  geom_point(
    data = xdata.agg %>% filter(Condition == "Functional"),
    aes(x = condition2),
    color = "dodgerblue",
    size = 1.5,
    alpha = .5
  ) +
  geom_point(
    data = xdata.agg %>% filter(Condition == "Non-functional"),
    aes(x = condition2),
    color = "darkorange",
    size = 1.5,
    alpha = .5,
    
  ) +
  geom_half_boxplot(
    data = xdata.agg %>% filter(Condition == "Functional"),
    aes(x = condition2, y = mean.peeking),
    position = position_nudge(x = -0.3),
    side = "l",
    outlier.shape = NA,
    center = TRUE,
    errorbar.draw = TRUE,
    width = .05,
    fill = 'dodgerblue',
    alpha = .5
  ) +
  
  geom_half_boxplot(
    data = xdata.agg %>% filter(Condition == "Non-functional"),
    aes(x = condition2, y = mean.peeking),
    position = position_nudge(x = 0.3),
    side = "r",
    outlier.shape = NA,
    center = TRUE,
    errorbar.draw = TRUE,
    width = .05,
    fill = 'darkorange',
    alpha = .5
  ) +
  
  # Define additional settings
  xlab("") +
  ylab("Mean peeking duration (s)") +
  scale_x_continuous(
    breaks = c(1, 2),
    labels = c("Functional", "Non-functional"),
    limits = c(0.5, 2.5)
  ) +
  #ylim(0, 1) +
  theme_classic()

sum_peeking_plot 

ggsave(sum_peeking_plot , filename = "Graphics/sum_peeking_plot.png", width = 6, height = 6, scale = 0.6)

```
## Plot: Mean proportion peeking duration
```{r}
#xdata.agg$condition<-fct_relevel(as.factor(xdata.agg$condition), "blocked", "teasing", "clumsy")
xdata.agg$condition2 <- jitter(as.numeric(as.factor(xdata.agg$Condition), amount = .0001))

library(gghalves)


sum_prop_peeking_plot <-
  ggplot(data = xdata.agg, aes(x = Condition, y = mean.prop.peeking, group = Condition)) +
  geom_line(
    aes(x = condition2, group = Subject),
    color = "darkgray",
    lty = 1,
    alpha = .3
  ) +
  geom_point(
    data = xdata.agg %>% filter(Condition == "Functional"),
    aes(x = condition2),
    color = "dodgerblue",
    size = 1.5,
    alpha = .5
  ) +
  geom_point(
    data = xdata.agg %>% filter(Condition == "Non-functional"),
    aes(x = condition2),
    color = "darkorange",
    size = 1.5,
    alpha = .5,
    
  ) +
  geom_half_boxplot(
    data = xdata.agg %>% filter(Condition == "Functional"),
    aes(x = condition2, y = mean.prop.peeking),
    position = position_nudge(x = -0.3),
    side = "l",
    outlier.shape = NA,
    center = TRUE,
    errorbar.draw = TRUE,
    width = .05,
    fill = 'dodgerblue',
    alpha = .5
  ) +
  
  geom_half_boxplot(
    data = xdata.agg %>% filter(Condition == "Non-functional"),
    aes(x = condition2, y = mean.prop.peeking),
    position = position_nudge(x = 0.3),
    side = "r",
    outlier.shape = NA,
    center = TRUE,
    errorbar.draw = TRUE,
    width = .05,
    fill = 'darkorange',
    alpha = .5
  ) +
  
  # Define additional settings
  xlab("") +
  ylab("Mean proportion peeking duration") +
  scale_x_continuous(
    breaks = c(1, 2),
    labels = c("Functional", "Non-functional"),
    limits = c(0.5, 2.5)
  ) +
  #ylim(0, 1) +
  theme_classic()

sum_prop_peeking_plot 

ggsave(sum_prop_peeking_plot , filename = "Graphics/kea_sum_prop_peeking_plot.png", width = 6, height = 6, scale = 0.6)

```

## Plot of peeking count
```{r}
xdata.agg2$condition2 <- jitter(as.numeric(as.factor(xdata.agg2$Condition), amount = .0001))

library(gghalves)


plot_binary_peeking_boxplot <-
  ggplot(data = xdata.agg2, aes(x = Condition, y = count, group = Condition)) +
  geom_line(
    aes(x = condition2, group = Subject),
    color = "darkgray",
    lty = 1,
    alpha = .3
  ) +
  geom_point(
    data = xdata.agg2 %>% filter(Condition == "Functional"),
    aes(x = condition2),
    color = "dodgerblue",
    size = 1.5,
    alpha = .5
  ) +
  geom_point(
    data = xdata.agg2 %>% filter(Condition == "Non-functional"),
    aes(x = condition2),
    color = "darkorange",
    size = 1.5,
    alpha = .5,
   
  ) +
  geom_half_boxplot(
    data = xdata.agg2 %>% filter(Condition == "Functional"),
    aes(x = condition2, y = count),
    position = position_nudge(x = -0.3),
    side = "l",
    outlier.shape = NA,
    center = TRUE,
    errorbar.draw = TRUE,
    width = .05,
    fill = 'dodgerblue',
    alpha = .5
  ) +
  
  geom_half_boxplot(
    data = xdata.agg2 %>% filter(Condition == "Non-functional"),
    aes(x = condition2, y = count),
    position = position_nudge(x = 0.3),
    side = "r",
    outlier.shape = NA,
    center = TRUE,
    errorbar.draw = TRUE,
    width = .05,
    fill = 'darkorange',
    alpha = .5
  ) +
  
  # Define additional settings
  xlab("") +
  ylab("Sum of trials with peeking") +
  scale_x_continuous(
    breaks = c(1, 2),
    labels = c("Functional", "Non-functional"),
    limits = c(0.5, 2.5)
  ) +
  #ylim(0, 1) +
  theme_classic()

plot_binary_peeking_boxplot
```
```{r}
ggsave(plot_binary_peeking_boxplot , filename = "Graphics/kea_binary_peeking_summed.png", width = 6, height = 6, scale = 0.7)

```
## Plot of peeking type by session
```{r}
xdata.agg.session<-xdata %>%
  group_by(Condition, Session) %>%
  summarise(mean.peeking = mean(Proportionpeeking), count = sum(Proportionpeeking > 0))

plot_binary_peeking_over_sessions<-ggplot(data = xdata.agg.session, aes(x = as.factor(Session), y = count)) +
  geom_bar(aes(fill=Condition), stat="identity", position = position_dodge(), color = "black")+
  theme_classic()+
  scale_fill_manual(values=c("Non-functional" = "darkorange", "Functional" = "dodgerblue"))+
  theme(legend.title = element_blank(), legend.position = c(0.82, 0.9))+
  labs(x = "Session", y = "Sum of peeking individuals")+
  scale_y_continuous(breaks = seq(0, 12, 1))
plot_binary_peeking_over_sessions
```

```{r}
ggsave(plot_binary_peeking_over_sessions , filename = "Graphics/kea_binary_peeking_count_over_session.png", width = 8, height = 6, scale = 0.6)
```

## Plot of peeking type
```{r}
peekingdata.agg<-peekingdata %>%
  group_by(Subject, Condition, Behavior) %>%
  summarise(mean.peeking = mean(sum_peeking))%>%
  mutate(Behavior = fct_recode(Behavior, "Time Spent on Right Side" = "Right box (side)"))

peekingdataplot<-ggplot(data = peekingdata.agg, aes(x = Condition, y = mean.peeking)) +
  geom_boxplot(outlier.colour = "white")+ 
  geom_jitter(width=0.3)+
  theme_bw()+
  facet_wrap(~Behavior, scales = "free_y")+
  ylab("Average Peeking Behavior")+
  xlab("")

peekingdataplot
ggsave(peekingdataplot, file="Graphics/Overall Peeking Data Plot.png", width = 10, height = 10, scale = 0.6)
```
# Analysis all data
## z-transformation of covariates (standardizes data)
Centering (subtract mean) and scaling (divide by SD so it becomes 1) covariates. Makes it more likely model will converge (does not change p value).
```{r}
xdata$z.trial <- as.vector(scale(as.numeric(xdata$Trial), scale = TRUE, center = TRUE))
xdata$z.session <- as.vector(scale(as.numeric(xdata$Session)))
```

## Dummy code factors (to prepare for centering below)
makes first factor 0, second factor 1
```{r}
xdata$Condition <- as.factor(xdata$Condition)
xdata$condition.dummy <- as.numeric(xdata$Condition == levels(xdata$Condition)[2])
```

## Center condition for random slopes (to avoid arbitrary effects of making first factor 0 and second factor 1)
```{r}
xdata$condition.c <- as.numeric(xdata$Condition) - mean(as.numeric(xdata$Condition))
mean(xdata$condition.c)
```

## Define control structure to make convergence more likely:
```{r}
contr <- glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000))
```

## Analysis of binary peeking response (overall peeks, includes time spent on right side)
```{r}
mm_peeking_overall_model<-glmer(
  peeking_binary ~ Condition + z.trial + z.session + (1 + condition.c + z.trial + z.session | Subject),
  data = xdata,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000))
)

drop1(mm_peeking_overall_model, test = "Chisq")
summary(mm_peeking_overall_model)
```


```{r}

mm1_peeking_overall_drop1 <- drop1(mm_peeking_overall_model, test="Chisq")%>% 
  filter(!is.na(npar)) %>% 
  add_row(npar = rep(NA,1),  .before = 1) 
```

Check for colinearity
```{r}
library(car)
xx=lm(peeking_binary ~ Condition + z.trial + z.session, data=xdata)
vif(xx)
```

```{r}
mm_peeking_overall_model.ci=boot.glmm.pred(model.res=mm_peeking_overall_model, excl.warnings=F,
nboots=1000, para=T, n.cores="all-1", resol=1000, level=0.95)

```


#### output table


```{r}
mm1_overall_peeking_output_table <-
  bind_cols(as.data.frame(summary(mm_peeking_overall_model)$coeff),
            mm1_peeking_overall_drop1,
            mm_peeking_overall_model.ci$ci.estimates) %>%
  select(
    Estimate,
    SE = `Std. Error`,
    LowerCI = X2.5.,
    UpperCI = X97.5.,
    Chi2 = LRT,
    df = npar,
    p = `Pr(Chi)`,
    z_wald = `z value`,
    p_wald = `Pr(>|z|)`
  ) %>% #
  mutate(across(.cols = c(p), ~ format(round(.x, 3), nsmall = 3))) %>%
  mutate(across(.cols = c(Estimate:Chi2), ~ format(round(.x, 2), nsmall = 2))) %>%
  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
  mutate(p = replace(p, p == 0, "<0.001"))

write.csv(mm1_overall_peeking_output_table, file = "saves/mm1_overall_peeking_output_table.csv")
```








## Right side peek (includes time spent on right side) analysis
```{r}
rightsidepeekdata <- peekingdata%>%
  filter(Behavior == "Right side peek")
```

## prepares for a plot of proportion peeking on the right side, looking at just subject and condition
```{r}
xdata.agg3<-peekingdata %>%
  filter(Behavior == "Right box (side)")%>%
  group_by(Subject, Condition) %>%
  summarise(mean.peeking = mean(Proportionpeeking), count = sum(Proportionpeeking > 0))

write.csv(xdata.agg3, file = "Saves/Time Spent Right Side Table.csv")
```

## Z-transformation of covariates (standardize subset of data, so have to transform again, same for dummy codes too)
```{r}
rightsidepeekdata$z.trial <- as.vector(scale(as.numeric(rightsidepeekdata$Trial)))
rightsidepeekdata$z.session <- as.vector(scale(as.numeric(rightsidepeekdata$Session)))
```

## Dummy code factors
```{r}
rightsidepeekdata$Condition <- as.factor(rightsidepeekdata$Condition)
rightsidepeekdata$condition.dummy <- as.numeric(rightsidepeekdata$Condition == levels(rightsidepeekdata$Condition)[2])
```

## Center condition for random slopes
```{r}
rightsidepeekdata$condition.c <- as.numeric(rightsidepeekdata$Condition) - mean(as.numeric(rightsidepeekdata$Condition))
```

## Define control structure to make convergence more likely
```{r}
contr <- glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000))
```

## Analysis of binary peeking response of just right side peek
```{r}
peekingmodel<-glmer(
  peeking_binary ~ Condition + z.trial + z.session + (1 + condition.c + z.trial + z.session | Subject),
  data = rightsidepeekdata,
  family = binomial,
  control = contr
)
```

## gives another way to evaluate the data, checks predictors
```{r}
drop1(peekingmodel, test = "Chisq")
summary(peekingmodel)
```



## Analysis of proportional peeking response
```{r}
library(glmmTMB)
library(car)

source("./functions/diagnostic_fcns.r")
source("./functions/glmm_stability.r")
source("./functions/boot_glmmTMB.r")
source("./functions/glmmTMB_stability.r")
source("./functions/drop1_para_glmmtmb.r")
source("./functions/extract_ranef_gmmTMB.r")
```

## Transformation check of ALL PEEKS proportion 
```{r}
contr<-glmmTMBControl(optCtrl=list(iter.max=100000000, eval.max=100000000))
xdata$Proportionpeeking_scaled <- (xdata$Proportionpeeking*(length(xdata$Proportionpeeking) - 1) + 0.5)/length(xdata$Proportionpeeking)
```

## check that ALL PEEKS proportional data is within bounds of 0 and 1 (and doesn't include 0s and 1s)
```{r}
min(xdata$Proportionpeeking_scaled)
max(xdata$Proportionpeeking_scaled)
max(xdata$Proportionpeeking)
```

## Analysis of proportional ALL peeking response
```{r}
peekingmodel_beta<-glmmTMB(
  Proportionpeeking_scaled ~ Condition + z.trial + z.session + 
    (1| Subject)+
    (0 + condition.c | Subject)+(0 +z.trial | Subject)+(0 +  z.session | Subject),
  data = xdata,
  family = beta_family,
  control = contr
)
```

## check if model is overdispersed or not (assumption check for proportion models)
```{r}
overdisp.test(peekingmodel_beta)
```


```{r}
summary(peekingmodel_beta)

drop1(peekingmodel_beta, test="Chisq")
```
LRT
```{r}
peekingmodel_beta_drop1 <- drop1(peekingmodel_beta, test="Chisq")%>% 
  filter(!is.na(Df)) %>% 
  add_row(Df = rep(NA,1),  .before = 1) 
```
 

+ Collinearity checks
```{r}
library(car)
xx=lm(Proportionpeeking_scaled ~ Condition + z.trial + z.session, data=xdata)
vif(xx)
```
confidence intervals
```{r eval=FALSE}
peekingmodel_beta.ci=boot.glmmtmb(peekingmodel_beta, 
nboots=1000, para=T, n.cores="all-1", resol=1000, level=0.95, data=xdata)
```

#### output table

```{r}
peekingmodel_beta_table <- bind_cols(as.data.frame(summary(peekingmodel_beta)$coefficients$cond),
                                      peekingmodel_beta_drop1)%>%#,
                                    # mm1_area_tail.ci$ci.estimates$fe[1:7,]) %>%
  dplyr::select(Estimate, SE = `Std. Error`, Chi2 = LRT, df = Df, p = `Pr(>Chi)`, z_wald=`z value`, p_wald=`Pr(>|z|)`) %>% #LowerCI = X2.5., UpperCI = X97.5., 
  mutate(across(.cols = c(p), ~ format(round(.x, 3), nsmall=3))) %>% 
  mutate(across(.cols = c(Estimate:Chi2), ~ format(round(.x, 2), nsmall=2))) %>% 
  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
  mutate(p=replace(p, p==0, "<0.001"))

write.csv(peekingmodel_beta_table , file = "saves/peekingmodel_beta_table.csv")
```






## Transformation check of RIGHT SIDE peek proportion
```{r}
rightsidepeekdata$Proportionpeeking_scaled <- (rightsidepeekdata$Proportionpeeking*(length(rightsidepeekdata$Proportionpeeking) - 1) + 0.5)/length(rightsidepeekdata$Proportionpeeking)
```

## check that RIGHT SIDE peek proportional data is within bounds of 0 and 1 (and doesn't include 0s and 1s)
```{r}
min(rightsidepeekdata$Proportionpeeking_scaled)
max(rightsidepeekdata$Proportionpeeking_scaled)
max(rightsidepeekdata$Proportionpeeking)
```

## Analysis of RIGHT SIDE peek proportion
```{r}
rightsidepeekmodel_beta<-glmmTMB(
  Proportionpeeking_scaled ~ Condition + z.trial + z.session + 
    (1| Subject)+
    (0 + condition.c | Subject)+(0 +z.trial | Subject)+(0 +  z.session | Subject),
  data = rightsidepeekdata,
  family = beta_family,
  control = contr
)
```

## check if model is overdispersed or not (assumption check for proportion models)
```{r}
overdisp.test(rightsidepeekmodel_beta)

summary(rightsidepeekmodel_beta)

drop1(rightsidepeekmodel_beta, test="Chisq")

```




# first session analysis

```{r}
first_session_data <- xdata %>%
  filter(Session == 1)

```

```{r}
xdata_session <- xdata %>%
  group_by(Condition, Session) %>%
  summarise(mean_prop_peeking = mean(Proportionpeeking), se = sd(Proportionpeeking) / sqrt(length(Proportionpeeking)))

peeking_session_plot <-
  ggplot(xdata_session,
         aes(
           x = as.factor(Session),
           y = mean_prop_peeking,
           fill = Condition
         )) +
  geom_bar(
    data = xdata_session,
    aes(
      x = as.factor(Session),
      y = mean_prop_peeking,
      group = interaction(Session, Condition),
      fill = Condition
    ),
    stat = "identity",
    position = position_dodge(width = 0.9)
  ) +
  geom_errorbar(
    data = xdata_session,
    aes(
      x = as.factor(Session),
      ymin = mean_prop_peeking - se,
      ymax = mean_prop_peeking + se
    ),
    position = position_dodge(width = 1),
    width = 0.4
  ) +
  
  #geom_boxplot(data = xdata, aes(x = Session, y = Proportionpeeking, group=interaction(Session, Condition), fill=Condition), position = position_dodge(width = .9))+
  #geom_count(position = position_dodge(width = .9), alpha=0.3)+
  # geom_line(aes(group = Condition)) +
  theme_bw() +
  ylim(0, 0.20) +
  scale_fill_manual(values = c("dodgerblue", "darkorange")) +
  ylab("Average Proportion Peeking Behavior") +
  xlab("Session")
#facet_wrap(~Condition)

ggsave(peeking_session_plot, file = "Graphics/peeking_session_plot.png", width = 9, height= 5, scale=0.65)
```



## Analysis of binary peeking response (overall peeks, includes time spent on right side) in first session
```{r}
mm_peeking_overall_model_session1<-glmer(
  peeking_binary ~ Condition + z.trial (1 + condition.c  | Subject),
  data = first_session_data,
  family = binomial,
  control = glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000))
)

drop1(mm_peeking_overall_model_session1, test = "Chisq")
summary(mm_peeking_overall_model_session1)
```





# Analysis of latency to first peek
```{r}



latency.data<-raw.data.after.solution %>%
  filter(Behavioral.category == "Peeking" | (Behavioral.category == "Proximity" & Behavior == "Right box (side)") | Behavior =="Food to end of trial") %>%
  group_by(Subject, Session, Trial, Condition, Behavior)%>% #Subject.1 means trial type (func or non func)
  summarise(first_instance=min(Start..s.)) %>%
  pivot_wider(names_from=Behavior, values_from = first_instance) %>%
  rename(solution = "Food to end of trial", right_side_peek = "Right side peek", right_box_side = "Right box (side)", peephole_peek = "Peephole peek", peek_on_top = "Peek on top")%>%
 # mutate(first_peek = pmap(across("Right side peek": "Peek on top"), min(na.rm=TRUE)))
  rowwise() %>%
  mutate(
    first_peek = min(c(right_side_peek, right_box_side, peephole_peek, peek_on_top), na.rm=TRUE)
  ) %>%
  ungroup() %>%
  mutate(first_peek = as.numeric(ifelse(first_peek==Inf, NA, first_peek)),
         latency_first_peek = first_peek - solution)

```

### fit LMM on latency data
#### Z-transformation of covariates (standardize subset of data, so have to transform again, same for dummy codes too)
```{r}
latency.data$z.trial <- as.vector(scale(as.numeric(latency.data$Trial)))
latency.data$z.session <- as.vector(scale(as.numeric(latency.data$Session)))
```

## Dummy code factors
```{r}
latency.data$Condition <- as.factor(latency.data$Condition)
latency.data$condition.dummy <- as.numeric(latency.data$Condition == levels(latency.data$Condition)[2])
```

## Center condition for random slopes
```{r}
latency.data$condition.c <- as.numeric(latency.data$Condition) - mean(as.numeric(latency.data$Condition))
```

```{r}
hist(latency.data$latency_first_peek)
```


fit model
```{r}

mm.latency<-lme4::lmer(latency_first_peek ~ Condition + z.trial + z.session + 
                         (1 + condition.c + z.trial + z.session | Subject),
                          data=latency.data, REML=FALSE)

summary(mm.latency)
```


```{r}
diagnostics.plot(mm.latency, size.fac=2)

ranef.diagn.plot(mm.latency)
```

```{r}
drop1_mm.latency <- drop1(mm.latency, test="Chisq")%>%
  filter(!is.na(npar))%>%
  add_row(npar = rep(NA,1),  .before = 1)

drop1_mm.latency

```

Colinearity checks
```{r}
library(car)
xx=lm(latency_first_peek ~ Condition + z.trial + z.session, data=latency.data)
vif(xx)
```
relative model complexity
```{r}
length(residuals(mm.latency))/
(length(fixef(mm.latency))+
nrow(as.data.frame(summary(mm.latency)$varcor)))
```
model stability
```{r}

mm.latency.stab=glmm.model.stab(model.res=mm.latency, contr=NULL, para=F, data=NULL)

mm.latency.stab$summary

m.stab.plot(round(mm.latency.stab$summary[, -1], 3))
```
--> model stable with regard to the fixed effects

```{r}
boot.mm.latency=boot.lmer(mm.latency, 
nboots=1000, para=T, n.cores="all-1", resol=1000, level=0.95)

boot.mm.latency_ci<-boot.mm.latency$ci.estimates
```
#### effect size
```{r}
library(MuMIn)
r.squaredGLMM(boot.mm.latency)
```


#### output table
```{r}
mm_latency_output_table <- bind_cols(as.data.frame(summary(mm.latency)$coefficients),
                         drop1_mm.latency) %>% #,
                         #boot.mm.latency_ci) %>% 
  dplyr::select(Estimate, SE = `Std. Error`,  Chi2 = LRT, df = npar, p = `Pr(Chi)`) %>% #LowerCI = X2.5., UpperCI = X97.5.,
  mutate(across(.cols = c(p), ~ round(.x, 3))) %>% 
  mutate(across(.cols = c(Estimate:Chi2), ~ round(.x, 2))) %>% 
  #mutate(across(Chi2:p, ~replace_na(.x, "")))%>%
mutate(p=replace(p, p==0, "<0.001"))

write.csv(mm_latency_output_table , file = "saves/mm_latency_output_table.csv")
```


```{r}
latency.agg.data <- latency.data %>%
  group_by(Subject, Condition) %>%
  summarise(mean_latency = mean(latency_first_peek, na.rm=TRUE))

table(latency.data$Subject, is.na(latency.data$latency_first_peek))
```



```{r}
ggplot(latency.agg.data, aes(x = Condition, y = mean_latency)) +
  geom_boxplot()+
  geom_point(alpha=0.3) +
  theme_bw()
```

```{r}
latency.agg.data$condition2 <- jitter(as.numeric(as.factor(latency.agg.data$Condition), amount = .00001))

library(gghalves)
library(ggsignif)


latency_plot <- ggplot(data = latency.agg.data, aes(x = Condition, y= mean_latency, group=Condition)) +
      geom_line(aes(x = condition2, group = Subject), color = "darkgray", lty = 1, alpha = .3) +
  
  geom_point(data = latency.agg.data %>% filter(Condition == "Functional"), aes(x = condition2), color = "dodgerblue", size = 1.5, alpha = .5, ) +
    geom_point(data = latency.agg.data %>% filter(Condition == "Non-functional"), aes(x = condition2), color = "darkorange", size = 1.5, alpha = .5, ) +

   geom_half_boxplot(
     data = latency.agg.data %>% filter(Condition == "Functional"), aes(x = condition2, y = mean_latency), position = position_nudge(x = -0.25), 
     side = "l",outlier.shape = NA, center = TRUE, errorbar.draw = TRUE, width = 0.1, 
     fill = 'dodgerblue', alpha = .5) +
  
     geom_half_boxplot(
     data = latency.agg.data %>% filter(Condition == "Non-functional"), aes(x = condition2, y = mean_latency), position = position_nudge(x = 0.25), 
     side = "r",outlier.shape = NA, center = TRUE, errorbar.draw = TRUE, width = 0.1, 
     fill = 'darkorange', alpha = .5) +
  # Define additional settings
  xlab("") +
  ylab("Mean latency to first peek (s)") +
  scale_x_continuous(breaks = c(1, 2), labels = c("Functional", "Non-functional"), limits = c(0.55,2.45)) +
 # ylim(0, 1) +
  theme_classic()

latency_plot 

ggsave(latency_plot , file = "Graphics/latency_plot.png", width = 6, height = 6, scale=0.6)
```

```{r}
ggplot(latency.data%>%filter(Subject=="Paul" | Subject=="John" | Subject=="Kermit" | Subject=="Pick" | Subject=="Frowin"), aes(x = Condition, y = latency_first_peek)) +
  geom_boxplot()+
  geom_point(alpha=0.3) +
  theme_bw()
```
```{r}
ggplot(xdata, aes(x = Condition, y = sum_peeking)) +
  geom_boxplot()+
  geom_point(alpha=0.3) +
  theme_bw()+
  facet_wrap(~Subject)
```

```{r}
latency.plot<-ggplot(latency.agg.data, aes(x = Condition, y = mean_latency)) +
  geom_boxplot()+
  geom_point(alpha=0.3) +
  theme_bw()

ggplot(latency.agg.data%>%filter(Subject=="Paul" | Subject=="John" | Subject=="Kermit" | Subject=="Pick" | Subject=="Frowin"), aes(x = Condition, y = mean_latency)) +
  geom_boxplot()+
  geom_point(alpha=0.3) +
  geom_line(aes(group=Subject))+
  theme_bw()
```

combine plots
```{r}
library(cowplot)
pg<-plot_grid(sum_peeking_plot, latency.plot, labels=c("A", "B"))
ggsave(pg, file="Graphics/pg_example.png", width = 10, height=5, scale=0.6)
```

```{r}
library(cowplot)
pg<-plot_grid(sum_peeking_plot, latency_plot, labels=c("A", "B"))
ggsave(pg, file="Graphics/pg_poster.png", width = 20, height=10, scale=0.3, dpi=1000)
```

