---
title: "Kea Explanation Seeking Analysis Rmd"
output:
  pdf_document: default
  html_document: default
date: "2023-02-07"
---
## Libraries
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lme4)
library(tidyverse)
library(summarytools)


source("./functions/diagnostic_fcns.r")
source("./functions/glmm_stability.r")
source("./functions/boot_glmm.r")
source("./functions/boot_glmm2.r")
source("./functions/glmmTMB_stability.r")
source("./functions/drop1_para_glmmtmb.r")
source("./functions/extract_ranef_gmmTMB.r")
```

# original
## Read in data
```{r}
raw.data<-read.csv("data/ALL SUBJECTS Explanation Seeking Coding 17.1.2023 csv.csv")
cb.data<-read.csv("data/counter balancing trials.csv")%>%
  select(-Date)%>%
  filter(Condition != "Refresher")
cb.data2<-expand.grid(Subject=levels(as.factor(cb.data$Subject)), Trial=c(4,5), Session=c(1:8), Behavior=c("Right side peek", "Peek on top", "Peephole peek"))
```

## Calculate duration of each trial
```{r}
trialdurationdata<-raw.data %>%
  filter(Behavior == "Food to end of trial") %>%
  rename(Condition = Subject.1, Trial_Duration = Duration..s.) %>%
  mutate(Trial = as.numeric(ifelse(Trial.Order == 1 & Condition == "Functional", 4, 
                                   ifelse(Trial.Order == 2 & Condition == "Functional", 5,
                                          ifelse(Trial.Order == 1 & Condition == "Non-functional", 5,
                                                 ifelse(Trial.Order == 2 & Condition == "Non-functional", 4,""))))))%>%
  select(Subject, Condition, Session, Trial,  Trial_Duration)

table(trialdurationdata$Subject,trialdurationdata$Condition)
table(cb.data$Subject,cb.data$Condition)
```
## Subset the raw data to the exploration period
first we determine when the food hit the ground and then we filter the data such that we only keep the peeks following the food release
```{r}
solution.data<-raw.data %>%
  filter( Behavior =="Food to end of trial") %>%
  rename(Condition = Subject.1) %>%
  mutate(Trial = as.numeric(ifelse(Trial.Order == 1 & Condition == "Functional", 4, 
                        ifelse(Trial.Order == 2 & Condition == "Functional", 5,
                               ifelse(Trial.Order == 1 & Condition == "Non-functional", 5,
                                      ifelse(Trial.Order == 2 & Condition == "Non-functional", 4,""))))))%>%
  group_by(Subject, Session, Trial, Condition)%>% #Subject.1 means trial type (func or non func)
  summarise(solution_time=min(Start..s.))


raw.data.after.solution <- raw.data %>%
  rename(Condition = Subject.1) %>%
  mutate(Trial = as.numeric(ifelse(Trial.Order == 1 & Condition == "Functional", 4, 
                        ifelse(Trial.Order == 2 & Condition == "Functional", 5,
                               ifelse(Trial.Order == 1 & Condition == "Non-functional", 5,
                                      ifelse(Trial.Order == 2 & Condition == "Non-functional", 4,""))))))%>%
  full_join(solution.data) %>%
  filter(Start..s.>=solution_time)
```


## Data preparation (summed up peeks, trial duration, added 0s, proportion peeking)
```{r}
xdata<-raw.data.after.solution %>%
  filter(Behavioral.category == "Peeking" | (Behavioral.category == "Proximity" & Behavior == "Right box (side)")) %>%
  group_by(Subject, Session, Trial, Condition)%>% #Subject.1 means trial type (func or non func)
  summarise(sum_peeking=sum(Duration..s.))%>% #aggregating durations per behavioral category within each session
  ungroup()%>%
  full_join(cb.data)%>%
  complete(Subject, Session, Trial,  fill=list(sum_peeking=0))%>% #fill in 0s
  full_join(trialdurationdata)%>%
  mutate(Proportionpeeking = sum_peeking / Trial_Duration,
         peeking_binary= as.numeric(ifelse(sum_peeking > 0, 1, 0)))
```
# reli data
## Read in data
```{r}
reli_data <- raw.data_reli %>%
  group_by(Subject, Session) %>%
  summarise(min(Date))%>%
  select(-"min(Date)")

```

```{r}
raw.data_reli<-read.csv("data/Reli SUBJECTS Explanation Seeking Reli Coding.csv")
```

## Calculate duration of each trial
```{r}
trialdurationdata_reli<-raw.data_reli %>%
  filter(Behavior == "Food to end of trial") %>%
  rename(Condition = Subject.1, Trial_Duration = Duration..s.) %>%
  mutate(Trial = as.numeric(ifelse(Trial.Order == 1 & Condition == "Functional", 4, 
                                   ifelse(Trial.Order == 2 & Condition == "Functional", 5,
                                          ifelse(Trial.Order == 1 & Condition == "Non-functional", 5,
                                                 ifelse(Trial.Order == 2 & Condition == "Non-functional", 4,""))))))%>%
  select(Subject, Condition, Session, Trial,  Trial_Duration)

```
## Subset the raw data to the exploration period
first we determine when the food hit the ground and then we filter the data such that we only keep the peeks following the food release
```{r}
solution.data_reli<-raw.data_reli %>%
  filter( Behavior =="Food to end of trial") %>%
  rename(Condition = Subject.1) %>%
  mutate(Trial = as.numeric(ifelse(Trial.Order == 1 & Condition == "Functional", 4, 
                        ifelse(Trial.Order == 2 & Condition == "Functional", 5,
                               ifelse(Trial.Order == 1 & Condition == "Non-functional", 5,
                                      ifelse(Trial.Order == 2 & Condition == "Non-functional", 4,""))))))%>%
  group_by(Subject, Session, Trial, Condition)%>% #Subject.1 means trial type (func or non func)
  summarise(solution_time=min(Start..s.))


raw.data.after.solution_reli <- raw.data_reli %>%
  rename(Condition = Subject.1) %>%
  mutate(Trial = as.numeric(ifelse(Trial.Order == 1 & Condition == "Functional", 4, 
                        ifelse(Trial.Order == 2 & Condition == "Functional", 5,
                               ifelse(Trial.Order == 1 & Condition == "Non-functional", 5,
                                      ifelse(Trial.Order == 2 & Condition == "Non-functional", 4,""))))))%>%
  full_join(solution.data_reli) %>%
  filter(Start..s.>=solution_time)
```


## Data preparation (summed up peeks, trial duration, added 0s, proportion peeking)
```{r}
xdata_reli<-raw.data.after.solution_reli %>%
  filter(Behavioral.category == "Peeking" | (Behavioral.category == "Proximity" & Behavior == "Right box (side)")) %>%
  group_by(Subject, Session, Trial, Condition)%>% #Subject.1 means trial type (func or non func)
  summarise(sum_peeking=sum(Duration..s.))%>% #aggregating durations per behavioral category within each session
  ungroup()%>%
  full_join(cb.data)%>%
  right_join(trialdurationdata_reli)%>%
  complete(Subject, Session, Trial,  fill=list(sum_peeking=0))%>% #fill in 0s
  inner_join(reli_data)%>%
  mutate(Proportionpeeking = sum_peeking / Trial_Duration,
         peeking_binary= as.numeric(ifelse(sum_peeking > 0, 1, 0)))%>%
  filter(!is.na(Condition))%>%
  select(Subject, Session, Trial, Condition, Prop_peeking_reli = Proportionpeeking, Peeking_binary_reli = peeking_binary)

```
# join original and reli data

```{r}

joined_reli_data <- xdata %>%
  full_join(xdata_reli)

write.csv(joined_reli_data, "data/joined_reli_data.csv")
```

```{r}
joined_reli_data<-read.csv("data/joined_reli_data.csv") %>%
  filter(!is.na(peeking_binary), !is.na(Peeking_binary_reli))
```

# Cohen's kappa
```{r}
library(psych)

#calculate Cohen's Kappa
cohen.kappa(x=cbind(joined_reli_data$peeking_binary,joined_reli_data$Peeking_binary_reli))


joined_reli_data %>% summarise(match = sum(peeking_binary == Peeking_binary_reli), non_match = sum(peeking_binary != Peeking_binary_reli))

```

#Intraclass correlation coefficients
```{r}
ICC(joined_reli_data%>%dplyr::select(Proportionpeeking, Prop_peeking_reli), missing=TRUE)
```



