---
title: "Kea explanation seeking - binomial GLMM - Power analysis"
author: "Christoph Völter"
date: "11/04/2022"
output: 
  html_document:
    theme: united
    toc: yes
    toc_depth: 4
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())
library(tidyverse)
library(cowplot)
library("gghalves")
library(ggthemes)

load("kea_explanation_seeking_sim_N16.RData")
```

Notes:
DV: peeking response (binary)
Design: 2 conditions (functional / intrinsic), within subject design
4 sessions, 4 trials per condition (increase if needed)
Prediction: peeking in 50% of intrinsic trials, and 25% in functioning trials
N = 16

## Generate data
```{r echo=FALSE, include=FALSE}
set.seed(100)
n.subject <- 16 # number subjects
n.per.subject <- 8 # observations per subject
n.per.condition <- 4 # observations per subject and condition
n.blocks <- 4
#age_range <- c(12:130) # age range between 1 and 13 years
fun.per<-c(0.15, 0.25) # functional condition
int.per<-c(0.4, 0.5) # intrinsic condition

subj.id <- as.factor(paste("subj", str_pad(1:n.subject, 2, pad = "0"), sep = "."))

#all order combinations
library(combinat)
con_order <- c("fun-int", "fun-int", "int-fun", "int-fun")
order<-unlist(permn(con_order))
temp.data<-as.data.frame(order)
temp.data$session<-as.vector(rep(1:4, 24))
temp.data <- temp.data %>%
  separate(order, c("1", "2"), sep = "-") %>%
  pivot_longer(cols="1":"2", names_to = "trial_w_session", values_to = "condition")

start.data <- as.data.frame(temp.data[1:128,]) #only keep data for 16 subjects
start.data$subj.id<-sort(rep(subj.id, n.per.subject)) 

start.data$trial <- rep(1:n.per.subject,n.subject)
#start.data$sex <- rep(c(rep("m", n.per.subject), rep("f", n.per.subject)), n.subject/2)

demo.data<-read.csv(file="data/demo_data.csv", sep=";")
start.data$sex <- rep(demo.data$sex,each=n.per.subject)
start.data$age <- rep(demo.data$age,each=n.per.subject)

# z-transformation of covariates
start.data$z.age <- as.vector(scale(start.data$age))

start.data$z.trial <- as.vector(scale(as.numeric(start.data$trial_w_session)))
start.data$z.session <- as.vector(scale(as.numeric(start.data$session)))

# dummy code factors
start.data$condition <- as.factor(start.data$condition)
start.data$condition.dummy <- as.numeric(start.data$condition == levels(start.data$condition)[2])


# center condition for random slopes:
start.data$condition.c <- as.numeric(start.data$condition) - mean(as.numeric(start.data$condition))

table(start.data$trial, start.data$condition)
table(start.data$sex, start.data$condition)
table(start.data$trial_w_session, start.data$condition)
```

## Simulation

```{r eval=FALSE, include=FALSE}
n.simus <- 1000
r.effects <- c(1.1) # random effects to be simulated
# with the intercept being -1.098612 (qlogis(0.25)) we assume a moderately large random intercept of 1.1.

r.slope.int <- c(1.1)
# with the estimate being -1.098612 (qlogis(0.25)-qlogis(0.5)) we assume a moderately large random slope of 1.1.

r.slope.trial <- 0.2
r.slope.session <- 0.2

# create object to store the simulation parameters and results:
all.res <- data.frame(expand.grid(
  n.per.subject = n.per.subject, r.effect = r.effects,
  r.slope.int = r.slope.int, r.slope.trial = r.slope.trial,
  r.slope.session = r.slope.session,
  fun.per = fun.per,
  int.per = int.per,
  simu = 1:n.simus
))
all.res$icpt <- NA
all.res$conditionint <- NA
all.res$re.sd <- NA
all.res$warns.full <- NA
all.res$warns.null <- NA
all.res$lrt.p.con <- NA
#all.res$lrt.p.age <- NA
all.res$full.null.p <- NA

all.ests <- matrix(NA, nrow = n.simus, ncol = 1)
colnames(all.ests) <- c("lrt.p.con")

# create data frame with design:
## done above

# load packages needed:
library(lme4)
# Loading required package: Matrix
library(kyotil) # we want to store info about convergence issues

# define control structure to make convergence more likely:
contr <- glmerControl(optimizer = "bobyqa", optCtrl = list(maxfun = 10000))

xdata <- start.data
m.mat <- model.matrix(object = ~condition + z.age + sex + z.trial + z.session , data = xdata) # create model martix

# run simulation
for (i in 1:nrow(all.res)) {

  set.seed(i) # allows to later replicate individual simulations

  # add age  (if it should be generated in each loop)
  #age <- sample(x = age_range, size = length(unique(xdata$subj.id)), replace = T)
  #xdata$age <- as.numeric(age[as.numeric(xdata$subj.id)])
  #xdata$z.age <- scale(xdata$age)
  #m.mat <- model.matrix(object = ~condition + z.age + sex + z.trial + z.session , data = xdata) # create model martix

  coefs <- c(
    "(Intercept)" = qlogis(all.res[i, "fun.per"]),
    "conditionint" = log(all.res[i, "int.per"] / (1 - all.res[i, "int.per"])) - log(all.res[i, "fun.per"] / (1 - all.res[i, "fun.per"])),
  "z.age" = 0,
  "sexm" = 0,
  "z.trial" = 0,
  "z.session" = 0
  )

  LP <- m.mat[, names(coefs)] %*% coefs # LP wrt fixed effects

  # add random effect to linear predictor:
  LP <- LP + rnorm(n = n.subject, sd = all.res[i, "r.effect"])[as.numeric(xdata$subj.id)] +
    rnorm(n = n.subject, sd = all.res[i, "r.slope.int"])[as.numeric(xdata$subj.id)] * xdata$condition.dummy +
    rnorm(n = n.subject, sd = all.res[i, "r.slope.trial"])[as.numeric(xdata$subj.id)] * xdata$z.trial +
    rnorm(n = n.subject, sd = all.res[i, "r.slope.session"])[as.numeric(xdata$subj.id)] * xdata$z.session 

  # generate response:
  xdata$correct <- rbinom(n = nrow(xdata), size = 1, prob = exp(LP) / (1 + exp(LP)))


  # fit full model:
  full <- keepWarnings(glmer(correct ~ condition + z.trial + z.session + (1 + condition.c + z.trial + z.session | subj.id),
    data = xdata, family = binomial, control = contr
  ))

  # store results:
  all.res[i, c("icpt", "conditionint", "z.trial", "z.session")] <- fixef(full$value)
  all.res[i, "re.sd"] <- as.data.frame(summary(full$value)$varcor)[1, "sdcor"]
  all.res[i, "warns.full"] <- nchar(paste(full$warnings, collapse = ""))
  all.res[i, "lrt.p.con"] <- as.data.frame(drop1(full$value, test = "Chisq"))["condition", "Pr(Chi)"]
    all.res[i, "lrt.trial.p"] <- as.data.frame(drop1(full$value, test = "Chisq"))["z.trial", "Pr(Chi)"]
    print(i)
}

save.image("kea_explanation_seeking_sim_N16.RData")
```

## Evaluation of results 

* number of warning per combinations of random effects (out of 1000 models per cell)  
Full model:  
```{r echo=FALSE}
#full model
tapply(X=all.res[, "warns.full"]>0, INDEX=all.res[, c("r.slope.int", "r.effect")],
FUN=sum)
#warning codes: 
#363: unable to evaluate scaled gradient. Model failed to converge: degenerate  Hessian with 1 negative eigenvalues
#205: Model is nearly unidentifiable: large eigenvalue ratio - Rescale variables?
```


## Only models that converged are evaluated from here on:  

```{r include=FALSE}
all.res2=subset(all.res, warns.full==0)
```


### How many models converged, have a significant full-null model comparison, and a significant LRT of condition?  
```{r echo=FALSE}
n.converged<- all.res2%>%
    group_by(int.per,fun.per, r.effect, r.slope.int ) %>%
  summarise(n.converged=length(lrt.p.con))


lrt.data2 <- all.res2 %>%
  #filter(full.null.p<0.05)%>%
  group_by(int.per, fun.per, r.effect, r.slope.int ) %>%
  summarise(lrt.p.con.median = median(lrt.p.con), 
            lrt.p.trial.median = median(lrt.trial.p),
            n.sign.lrt.con = length(lrt.p.con[lrt.p.con < 0.05]),
            n.sign.lrt.trial = length(lrt.trial.p[lrt.trial.p < 0.05]),
            n.lrt = n.simus,
            proportion.sign.lrt.con = length(lrt.p.con[lrt.p.con < 0.05]) / n.simus,
            proportion.sign.lrt.trial = length(lrt.trial.p[lrt.trial.p < 0.05]) / n.simus)%>%
  full_join(n.converged)

lrt.data2
```

#### Plotting the proportion of significant LRTs for the predictor variable condition ONLY based on models that converged and with a significant full-null model comparison

```{r echo=FALSE}
p.con.power_4tr <- ggplot(data = lrt.data2, aes(x= as.factor(int.per),y = proportion.sign.lrt.con, fill=as.factor(fun.per))) +
  geom_bar(stat="identity", color="black", position=position_dodge())+
  scale_y_continuous(breaks=seq(0,1,0.2), limits=c(0, 1))+
  geom_hline(yintercept = 0.8, colour = "black", lwd = 1, lty = 2) +
   # geom_hline(yintercept = 0.05, colour = "darkgrey", lwd = 1.1, lty = 4) +
  scale_fill_manual(values=c("dodgerblue", "darkorange", "darkgrey"))+
  labs(fill = "Functional condition", y="Power", x= "Intrinsic condition") +
  theme_few()+
  ggtitle("4 trials per condition")+
  theme(legend.position="none")+
    geom_text(aes(label=round(proportion.sign.lrt.con,2)), position=position_dodge(width=0.9), vjust=-0.25)
p.con.power_4tr

ggsave(p.con.power_4tr, filename = "graphics/kea_explanation_seeking_power_8trials_16subj.png", scale = 0.7, height = 5, width = 8)
```

#### Plotting the intercepts


```{r echo=FALSE}
ggplot(data = all.res2, aes(x=as.factor(r.effect), y=icpt))+
  geom_jitter( alpha=0.5, col="grey")+
  geom_boxplot(aes(x=as.factor(r.effect), y=icpt, group=r.effect), alpha=0.1, outlier.colour="white")+
  geom_hline(yintercept=coefs["(Intercept)"], colour="red", lty=2)
```



#### Plotting the fixed effect of condition

```{r echo=FALSE}
p.con <- ggplot(data = all.res2, aes(x= as.factor(int.per), fill=as.factor(fun.per)))  +
  geom_jitter(data = all.res2, aes(x = as.factor(int.per), y = conditionint, color = as.factor(fun.per)), size = 1.5, position = position_jitterdodge(dodge.width = 0.8, jitter.width = 0.5), alpha = .1) +
  scale_color_manual(values = c("dodgerblue", "darkorange")) +
    geom_boxplot(data = all.res2 %>% filter(fun.per == "0.15"), aes(x = as.factor(int.per), y = conditionint), position = position_nudge(x = -.2), width = 0.3, alpha = 0.15, outlier.colour = "white") +
  geom_boxplot(data = all.res2 %>% filter(fun.per == "0.25"), aes(x = as.factor(int.per), y = conditionint), position = position_nudge(x = .2), width = 0.3, alpha = 0.15, outlier.colour = "white") +
  geom_hline(data = data.frame(int.per = "0.5"), aes(yintercept = coefs["conditionint"]), colour = "black", lwd = 1.1, lty = 2, alpha = 0.7) +
  geom_hline(data = data.frame(int.per = "0.4"), aes(yintercept = qlogis(0.4) - qlogis(0.15)), colour = "darkgrey", lwd = 1.1, lty = 4, alpha = 0.7) +
  ylab("Condition (fixed effect)") +
  xlab("Intrinsic condition") +
  ylim(-10, 10)+
  theme_few() #+
  #theme(legend.position = "none")

p.con 
```


```{r echo=FALSE}
p.legend <- ggplot(data = lrt.data2, aes(x= as.factor(int.per),y = proportion.sign.lrt.con, fill=as.factor(fun.per))) +
  geom_bar(stat="identity", color="black", position=position_dodge())+
  scale_y_continuous(breaks=seq(0,1,0.2), limits=c(0, 1))+
  geom_hline(yintercept = 0.95, colour = "black", lwd = 1, lty = 2) +
    geom_hline(yintercept = 0.05, colour = "darkgrey", lwd = 1, lty = 4) +
  scale_fill_manual(values=c("dodgerblue", "darkorange"))+
  labs(fill = "Functional condition", y="Power", x= "Intrinsic condition") +
  theme_few()+
  theme(legend.position="top")

p.leg <- get_legend(p.legend)

p.con <- plot_grid(perf.plot, p.con.power, p.con.str, p.con.intr, labels = c("A", "B", "C", "D"), rel_widths = c(1, 1, 1, 1), nrow=2)
p.con2 <- ggdraw(plot_grid(p.con, p.leg, ncol = 1, nrow = 2, rel_heights = c(1, 0.1)))

p.con2
ggsave(p.con2, filename = "graphics/Explanation seeking_children simulation.png", scale = 0.65, height = 12, width = 12)
```
